### 2024-03-11 paper/paper.tex
Add paper.tex
To update the TeX file, run the following in the folder of the draft:
```bash
docker run --rm \
    --volume $PWD:/data \
    --user $(id -u):$(id -g) \
    --env JOURNAL=joss \
    openjournals/paperdraft -k paper.md
```

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks=true,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{paper.bib}

% --- Splitting \texttt --------------------------------------------------

\let\textttOrig=\texttt
\def\texttt#1{\expandafter\textttOrig{\seqsplit{#1}}}
\renewcommand{\seqinsert}{\ifmmode
  \allowbreak
  \else\penalty6000\hspace{0pt plus 0.02em}\fi}

% --- Pandoc does not distinguish between links like [foo](bar) and
% --- [foo](foo) -- a simplistic Markdown model.  However, this is
% --- wrong:  in links like [foo](foo) the text is the url, and must
% --- be split correspondingly.
% --- Here we detect links \href{foo}{foo}, and also links starting
% --- with https://doi.org, and use path-like splitting (but not
% --- escaping!) with these links.
% --- Another vile thing pandoc does is the different escaping of
% --- foo and bar.  This may confound our detection.
% --- This problem we do not try to solve at present, with the exception
% --- of doi-like urls, which we detect correctly.

\makeatletter
\let\href@Orig=\href
\def\href@Urllike#1#2{\href@Orig{#1}{\begingroup
    \def\Url@String{#2}\Url@FormatString
    \endgroup}}
\def\href@Notdoi#1#2{\def\tempa{#1}\def\tempb{#2}%
  \ifx\tempa\tempb\relax\href@Urllike{#1}{#2}\else
  \href@Orig{#1}{#2}\fi}
\def\href#1#2{%
  \IfBeginWith{#1}{https://doi.org}%
  {\href@Urllike{#1}{#2}}{\href@Notdoi{#1}{#2}}}
\makeatother

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=2.54cm, left=2.54cm,
            headheight=2.2cm]{geometry}

% --- Default font ------------------------------------------------------------
\renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.0pt}

\fancyfoot[L]{}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\renewcommand{\affil}[2][]%
   {\newaffiltrue\let\AB@blk@and\AB@pand
      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
        \setcounter{Maxaffil}{0}\fi
        \begingroup
        \let\href=\href@Orig
        \let\texttt=\textttOrig
        \let\protect\@unexpandable@protect
        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
        \@temptokena=\expandafter{\AB@authors}%
        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
         \gdef\AB@las{}\gdef\AB@au@str{}%
        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
        \@temptokena=\expandafter{\AB@affillist}%
        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
      \endgroup
       \let\AB@affilsep\AB@affilsepx
}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{fontspec}

  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={jax\_op: Use any function in JAX},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

% --- We redefined \texttt, but in sections and captions we want the
% --- old definition
\let\addcontentslineOrig=\addcontentsline
\def\addcontentsline#1#2#3{\bgroup
  \let\texttt=\textttOrig\addcontentslineOrig{#1}{#2}{#3}\egroup}
\let\markbothOrig\markboth
\def\markboth#1#2{\bgroup
  \let\texttt=\textttOrig\markbothOrig{#1}{#2}\egroup}
\let\markrightOrig\markright
\def\markright#1{\bgroup
  \let\texttt=\textttOrig\markrightOrig{#1}\egroup}

\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\title{\texttt{jax\_op}: Use any function in JAX}

        \author[1, 2, 3*]{Jakob Roth}
          \author[1*]{Martin Reinecke}
          \author[1, 2*]{Gordian Edenhofer}

      \affil[1]{Max Planck Institute for Astrophysics,
Karl-Schwarzschild-Str. 1, 85748 Garching, Germany}
      \affil[2]{Ludwig Maximilian University of Munich,
Geschwister-Scholl-Platz 1, 80539 Munich, Germany}
      \affil[3]{Technical University of Munich, Boltzmannstr. 3, 85748
Garching, Germany}
      \affil[*]{These authors contributed equally.}
  \date{\vspace{-7ex}}

\begin{document}
\maketitle

\vspace{1em}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

JAX is widely used in machine learning and scientific computing.
Scientific computing relies on existing high-performance code which we
would ideally like to use in JAX. Reimplementing the existing code in
JAX is often impractical and the existing interface in JAX for
connecting custom code requires deep knowledge of JAX and its C++
backend. The aim of \texttt{jax\_op} is to drastically lower the burden
of connecting custom functions implemented in other programming
languages to JAX. Specifically, \texttt{jax\_op} provides an easy-to-use
Python interface for defining custom, so-called JAX primitives
supporting any JAX transformations.

\hypertarget{statement-of-need}{%
\section{Statement of Need}\label{statement-of-need}}

The use of JAX (Bradbury et al., 2018) is widespread in the natural
sciences. JAX's powerful transformation system is of especially high
interest. It enables retrieving arbitrary derivatives of functions,
batch computations, and just-in-time code compilation for additional
performance. Its transformation system relies on all constituents of the
computation being written in JAX.

A plethora of high-performance code is not written in JAX and thus not
accessible from within JAX. Rewriting these is often infeasible and/or
inefficient. Ideally, we would like to intermix existing
high-performance code with JAX code. However, connecting code to JAX
requires knowledge of the internals of JAX and its C++ backend.

In this paper, we present \texttt{jax\_op}, a package for bridging any
function to JAX without in-depth knowledge of JAX's transformation
system. The interface is accessible from Python without requiring any
development in C++. The package is able to register any function, its
partial derivatives and their transpose functions as a JAX native call,
a so-called primitive. Derivatives, compilation rules, and batching
rules are automatically registered with JAX.

We believe \texttt{jax\_op} to be highly useful in scientific computing.
We intend to use this package to connect the Hartley transform and the
spherical harmonic transform from ducc (Reinecke, 2024) to NIFTy
(Edenhofer et al., 2024) as well as the radio interferometry response
from ducc with \texttt{resolve} (Arras et al., 2024) for radio
astronomy. Furthermore, we intend to connect the non-uniform FFT from
ducc with JAX for applications in strong-lensing astrophysics. We
envision many further applications within and outside of astrophysics.

To the best of our knowledge no other code currently exists for
connecting generic functions to JAX. The package that comes the closest
is Enzyme-JAX (EnzymeAD, 2024). Enzyme-JAX allows one to differentiate a
C++ function with Enzyme (W. Moses \& Churavy, 2020; W. S. Moses et al.,
2021, 2022) and connect it together with its derivative to JAX. However,
it enforces the use of Enzyme for deriving derivatives and does not
allow for connecting arbitrary code to JAX.

\hypertarget{automatic-differentiation-and-code-example}{%
\section{Automatic Differentiation and Code
Example}\label{automatic-differentiation-and-code-example}}

Automatic differentiation is a core feature of JAX and often one of the
main reasons for using it. Thus, it is essential that custom functions
registered with JAX support automatic differentiation. In the following,
we will outline which functions our package respectively JAX requires to
enable automatic differentiation. For simplicity, we assume that we want
to connect the nonlinear function \(f(x_1,x_2) = x_1x_2^2\) to JAX. The
\texttt{jax\_op} package expects the Python function for \(f\) to take
three positional arguments. The first argument, \texttt{out}, is a
\texttt{tuple} into which the function results are written. The second
argument is also a \texttt{tuple} containing the input to the function,
in our case, \(x_1\) and \(x_2\). Via \texttt{kwargs\_dump}, potential
keyword arguments given to the later registered Jax primitive can be
forwarded to \texttt{f} in serialized form.

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{import}\NormalTok{ jax\_linop }\CommentTok{\#FIXME}

\KeywordTok{def}\NormalTok{ f(out, args, kwargs\_dump):}
\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jax\_linop.load\_kwargs(kwargs\_dump)}
\NormalTok{    x1, x2 }\OperatorTok{=}\NormalTok{ args}
\NormalTok{    out[}\DecValTok{0}\NormalTok{][()] }\OperatorTok{=}\NormalTok{ x1 }\OperatorTok{*}\NormalTok{ x2}\OperatorTok{**}\DecValTok{2}
\end{Highlighting}
\end{Shaded}

JAX's automatic differentiation engine can compute the Jacobian-vector
product \texttt{jvp} and vector-Jacobian product \texttt{vjp} of JAX
primitives. The Jacobian-vector product in JAX is a function applying
the Jacobian of \(f\) at a position \(x\) to a tangent vector. In
mathematical nomenclature this operation is called the pushforward of
\(f\) and can be denoted as \(\partial f(x): T_x X \mapsto T_{f(x)} Y\),
with \(T_x X\) and \(T_{f(x)} Y\) being the tangent spaces of \(X\) and
\(Y\) at the positions \(x\) and \(f(x)\). As the implementation of
\(f\) is not JAX native, JAX cannot automatically compute the
\texttt{jvp}. Instead, an implementation of the pushforward has to be
provided, which \texttt{jax\_op} will register as the \texttt{jvp} of
the JAX primitive of \(f\). For our example, this
Jacobian-vector-product function is given by
\(\partial f(x_1,x_2)(dx_1,dx_2) = x_2^2dx_1 + 2x_1x_2dx_2\).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{def}\NormalTok{ f\_jvp(out, args, kwargs\_dump):}
\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jax\_linop.load\_kwargs(kwargs\_dump)}
\NormalTok{    x1, x2, dx1, dx2 }\OperatorTok{=}\NormalTok{ args}
\NormalTok{    out[}\DecValTok{0}\NormalTok{][()] }\OperatorTok{=}\NormalTok{ x2}\OperatorTok{**}\DecValTok{2} \OperatorTok{*}\NormalTok{ dx1 }\OperatorTok{+} \DecValTok{2} \OperatorTok{*}\NormalTok{ x1 }\OperatorTok{*}\NormalTok{ x2 }\OperatorTok{*}\NormalTok{ dx2}
\end{Highlighting}
\end{Shaded}

The vector-Jacobian product \texttt{vjp} in JAX is the linear transpose
of the Jacobian-vector product. In mathematical nomenclature this is the
pullback \((\partial f(x))^{T}: T_{f(x)}Y \mapsto T_x X\) of \(f\).
Analogously to the \texttt{jvp}, the user has to implement this function
as JAX cannot automatically construct it. For our example function, the
vector-Jacobian product is
\((\partial f(x_1,x_2))^{T}(dy) = (x_2^2dy, 2x_1x_2dy)\).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{def}\NormalTok{ f\_vjp(out, args, kwargs\_dump):}
\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jax\_linop.load\_kwargs(kwargs\_dump)}
\NormalTok{    x1, x2, dy }\OperatorTok{=}\NormalTok{ args}
\NormalTok{    out[}\DecValTok{0}\NormalTok{][()] }\OperatorTok{=}\NormalTok{ x2}\OperatorTok{**}\DecValTok{2} \OperatorTok{*}\NormalTok{ dy}
\NormalTok{    out[}\DecValTok{1}\NormalTok{][()] }\OperatorTok{=} \DecValTok{2} \OperatorTok{*}\NormalTok{ x1 }\OperatorTok{*}\NormalTok{ x2 }\OperatorTok{*}\NormalTok{ dy}
\end{Highlighting}
\end{Shaded}

To just-in-time compile the function, JAX needs to abstractly evaluate
the code, i.e.~it needs to be able to know the shape and dtype of the
output of the custom function given only the shape and dtype of the
input. We have to provide these abstract evaluation functions returning
the output shape and dtype given an input shape and dtype for \texttt{f}
as well as for the \texttt{vjp} application. The output shape of the
\texttt{jvp} is identical to the output shape of \texttt{f} itself and
does not need to be specified again.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{def}\NormalTok{ f\_abstract(}\OperatorTok{*}\NormalTok{args, }\OperatorTok{**}\NormalTok{kwargs):}
    \ControlFlowTok{assert}\NormalTok{ args[}\DecValTok{0}\NormalTok{].shape }\OperatorTok{==}\NormalTok{ args[}\DecValTok{1}\NormalTok{].shape}
    \ControlFlowTok{return}\NormalTok{ ((args[}\DecValTok{0}\NormalTok{].shape, args[}\DecValTok{0}\NormalTok{].dtype),)}

\KeywordTok{def}\NormalTok{ f\_abstract\_T(}\OperatorTok{*}\NormalTok{args, }\OperatorTok{**}\NormalTok{kwargs):}
    \ControlFlowTok{return}\NormalTok{ (}
\NormalTok{        (args[}\DecValTok{0}\NormalTok{].shape, args[}\DecValTok{0}\NormalTok{].dtype),}
\NormalTok{        (args[}\DecValTok{0}\NormalTok{].shape, args[}\DecValTok{0}\NormalTok{].dtype),}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

We have now defined all ingredients necessary to register a JAX
primitive for our function \(f\) using the \texttt{jax\_op} package.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{f\_jax }\OperatorTok{=}\NormalTok{ jax\_linop.get\_nonlinear\_call(}
\NormalTok{    f, (f\_jvp, f\_vjp), f\_abstract, f\_abstract\_T}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\texttt{f\_jax} is a JAX primitive registered via the \texttt{jax\_op}
package supporting all JAX transformations. We can now compute the
\texttt{jvp} and \texttt{vjp} of the new JAX primitive and even
jit-compile and batch it.

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{import}\NormalTok{ jax}
\ImportTok{import}\NormalTok{ jax.numpy }\ImportTok{as}\NormalTok{ jnp}

\NormalTok{inp }\OperatorTok{=}\NormalTok{ (jnp.full((}\DecValTok{4}\NormalTok{,}\DecValTok{3}\NormalTok{), }\FloatTok{4.}\NormalTok{), jnp.full((}\DecValTok{4}\NormalTok{,}\DecValTok{3}\NormalTok{), }\FloatTok{2.}\NormalTok{))}
\NormalTok{tan }\OperatorTok{=}\NormalTok{ (jnp.full((}\DecValTok{4}\NormalTok{,}\DecValTok{3}\NormalTok{), }\FloatTok{1.}\NormalTok{), jnp.full((}\DecValTok{4}\NormalTok{,}\DecValTok{3}\NormalTok{), }\FloatTok{1.}\NormalTok{))}
\NormalTok{res, res\_tan }\OperatorTok{=}\NormalTok{ jax.jvp(f\_jax, inp, tan)}

\NormalTok{cotan }\OperatorTok{=}\NormalTok{ (jnp.full((}\DecValTok{4}\NormalTok{,}\DecValTok{3}\NormalTok{), }\FloatTok{6.}\NormalTok{),)}
\NormalTok{res, f\_vjp }\OperatorTok{=}\NormalTok{ jax.vjp(f\_jax, }\OperatorTok{*}\NormalTok{inp)}
\NormalTok{res\_cotan }\OperatorTok{=}\NormalTok{ f\_vjp(cotan)}

\NormalTok{f\_jax\_jit }\OperatorTok{=}\NormalTok{ jax.jit(f\_jax)}
\NormalTok{res }\OperatorTok{=}\NormalTok{ f\_jax\_jit(}\OperatorTok{*}\NormalTok{inp)}
\end{Highlighting}
\end{Shaded}

\hypertarget{higher-order-derivatives-and-linear-functions}{%
\section{Higher Order Derivatives and Linear
Functions}\label{higher-order-derivatives-and-linear-functions}}

JAX supports higher order derivatives and can differentiate a
\texttt{jvp} or \texttt{vjp} with respect to the position at which the
Jacobian was taken. Similar to first derivatives, JAX can not
automatically compute higher derivatives of a general function \(f\)
that is not natively implemented in JAX. Higher order derivatives would
again need to be provided by the user. For many algorithms, first
derivatives are sufficient, and higher order derivatives are often not
implemented by the high-performance codes. Therefore, the current
interface of \texttt{jax\_op} is, for simplicity, restricted to first
derivatives. In the future, the interface could be easily expanded if
specific use cases require higher order derivatives.

In scientific computing, linear functions such as, e.g., spherical
harmonic transforms are widespread. If the function \(f\) is linear,
differentiation becomes trivial. Specifically for a linear function
\(f\), the pushforward respectively the \texttt{jvp} of \(f\) is
identical to \(f\) itself and independent of the position at which it is
computed. Expressed in formulas, \(\partial f(x)(dx) = f(dx)\) if \(f\)
is linear in \(x\). Analogously, the pullback respectively the
\texttt{vjp} becomes independent of the initial position and is given by
the linear transpose of \(f\), thus
\((\partial f(x))^{T}(dy) = f^T(dy)\). Also, all higher order
derivatives can be expressed in terms of \(f\) and its transpose. To
make use of these simplifications, \texttt{jax\_op} provides a special
interface for linear functions, supporting higher order derivatives,
only requiring an implementation of the function and its transpose.

\hypertarget{platforms}{%
\section{Platforms}\label{platforms}}

Currently, \texttt{jax\_op} only has CPU but no GPU support. With some
expertise on Python bindings for GPU kernels adding GPU support should
be fairly simple. We especially want to highlight that the interfacing
with the JAX automatic differentiation engine is identical for CPU and
GPU.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We would like to thank Dan Foreman-Mackey for his detailed guide
(https://dfm.io/posts/extending-jax/) on connecting C++ code to JAX.
Jakob Roth acknowledges financial support from the German Federal
Ministry of Education and Research (BMBF) under grant 05A23WO1
(Verbundprojekt D-MeerKAT III). Gordian Edenhofer acknowledges support
from the German Academic Scholarship Foundation in the form of a PhD
scholarship (``Promotionsstipendium der Studienstiftung des Deutschen
Volkes'').

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Resolve2024}{}%
Arras, P., Roth, J., Ding, S., Reinecke, M., Fuchs, R., \& Johnson, V.
(2024). \emph{RESOLVE}. \url{https://gitlab.mpcdf.mpg.de/ift/resolve}

\leavevmode\hypertarget{ref-Jax2018}{}%
Bradbury, J., Frostig, R., Hawkins, P., Johnson, M. J., Leary, C.,
Maclaurin, D., Necula, G., Paszke, A., VanderPlas, J., Wanderman-Milne,
S., \& Zhang, Q. (2018). \emph{{JAX}: Composable transformations of
{P}ython+{N}um{P}y programs} (Version 0.3.13) {[}Computer software{]}.
\url{http://github.com/google/jax}

\leavevmode\hypertarget{ref-Edenhofer2023NIFTyRE}{}%
Edenhofer, G., Frank, P., Roth, J., Leike, R. H., Guerdi, M.,
Scheel-Platz, L. I., Guardiani, M., Eberle, V., Westerkamp, M., \&
Enßlin, T. A. (2024). \emph{{Re-Envisioning Numerical Information Field
Theory (NIFTy.re): A Library for Gaussian Processes and Variational
Inference}}. \url{http://arxiv.org/abs/2402.16683}

\leavevmode\hypertarget{ref-Moses2024}{}%
EnzymeAD. (2024). \emph{{Enzyme-JAX}} (Version 0.0.6) {[}Computer
software{]}. \url{https://github.com/EnzymeAD/Enzyme-JAX}

\leavevmode\hypertarget{ref-Moses2020}{}%
Moses, W., \& Churavy, V. (2020). Instead of rewriting foreign code for
machine learning, automatically synthesize fast gradients. In H.
Larochelle, M. Ranzato, R. Hadsell, M. F. Balcan, \& H. Lin (Eds.),
\emph{Advances in neural information processing systems} (Vol. 33, pp.
12472--12485). Curran Associates, Inc.
\url{https://proceedings.neurips.cc/paper/2020/file/9332c513ef44b682e9347822c2e457ac-Paper.pdf}

\leavevmode\hypertarget{ref-Moses2021}{}%
Moses, W. S., Churavy, V., Paehler, L., Hückelheim, J., Narayanan, S. H.
K., Schanen, M., \& Doerfert, J. (2021). Reverse-mode automatic
differentiation and optimization of GPU kernels via enzyme.
\emph{Proceedings of the International Conference for High Performance
Computing, Networking, Storage and Analysis}.
\url{https://doi.org/10.1145/3458817.3476165}

\leavevmode\hypertarget{ref-Moses2022}{}%
Moses, W. S., Narayanan, S. H. K., Paehler, L., Churavy, V., Schanen,
M., Hückelheim, J., Doerfert, J., \& Hovland, P. (2022). Scalable
automatic differentiation of multiple parallel paradigms through
compiler augmentation. \emph{Proceedings of the International Conference
on High Performance Computing, Networking, Storage and Analysis}.
ISBN:~\href{https://worldcat.org/isbn/9784665454445}{9784665454445}

\leavevmode\hypertarget{ref-ducc0}{}%
Reinecke, M. (2024). \emph{{DUCC}: Distinctly useful code collection}
(Version 0.33.0) {[}Computer software{]}.
\url{https://gitlab.mpcdf.mpg.de/mtr/ducc}

\end{CSLReferences}

\end{document}

### 2024-03-11 demos/demo_lin.py
jax_linop to jaxbind

@@ -5,7 +5,7 @@ import jax
 from jax import numpy as jnp
 from jax.test_util import check_grads
 
-import jax_linop
+import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
@@ -18,7 +18,7 @@ def lin(out, args, kwargs_dump):
     # %timeit pickle.loads(np.ndarray.tobytes(kwargs_dump))
     # # 582 ns ± 1.12 ns per loop (mean ± std. dev. of 7 runs, 1,000,000 loops each)
     # ```
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     x, y = args
     out[0][()] = x + y
     out[1][()] = x + y
@@ -44,7 +44,7 @@ def lin_abstract_T(*args, **kwargs):
     return ((a.shape, a.dtype), (a.shape, a.dtype))
 
 
-lin_jax = jax_linop.get_linear_call(
+lin_jax = jaxbind.get_linear_call(
     lin, lin_T, lin_abstract, lin_abstract_T, first_n_args_fixed=0,
 )
 inp = (4 + jnp.zeros((2, 2)), 1 + jnp.zeros((2, 2)))
@@ -79,7 +79,7 @@ def lin_abstract_T(*args, **kwargs):
     return ((x.shape, x.dtype),)
 
 
-lin_jax = jax_linop.get_linear_call(
+lin_jax = jaxbind.get_linear_call(
     lin,
     lin_T,
     lin_abstract,

### 2024-03-11 demos/demo_mlin.py

@@ -6,7 +6,7 @@ import numpy as np
 from jax import numpy as jnp
 from jax.test_util import check_grads
 
-import jax_linop
+import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
@@ -46,7 +46,7 @@ def mlin_abstract_T2(*args, **kwargs):
 
 func_T = (mlin_T1, mlin_T2)
 func_abstract_T = (mlin_abstract_T1, mlin_abstract_T2)
-mlin_jax = jax_linop.get_linear_call(
+mlin_jax = jaxbind.get_linear_call(
     mlin,
     func_T,
     mlin_abstract,
@@ -105,7 +105,7 @@ def mlin_abstract_T2(*args, **kwargs):
 
 func_T = (None, mlin_T2)
 func_abstract_T = (None, mlin_abstract_T2)
-mlin_jax = jax_linop.get_linear_call(
+mlin_jax = jaxbind.get_linear_call(
     mlin,
     func_T,
     mlin_abstract,

### 2024-03-11 demos/demo_mlin_batch.py

@@ -6,7 +6,7 @@ import numpy as np
 from jax import numpy as jnp
 from jax.test_util import check_grads
 
-import jax_linop
+import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
@@ -16,7 +16,7 @@ def mlin_call(x, y):
 
 
 def mlin(out, args, kwargs_dump):
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     batch_axes = kwargs.pop("batch_axes", None)
     call = mlin_call
     if batch_axes is not None and batch_axes != ((),) * len(args):
@@ -63,7 +63,7 @@ def mlin_abstract_T2(*args, **kwargs):
 
 func_T = (mlin_T1, mlin_T2)
 func_abstract_T = (mlin_abstract_T1, mlin_abstract_T2)
-mlin_jax = jax_linop.get_linear_call(
+mlin_jax = jaxbind.get_linear_call(
     mlin,
     func_T,
     mlin_abstract,

### 2024-03-11 demos/demo_nonlin.py

@@ -6,14 +6,14 @@ import numpy as np
 from jax import numpy as jnp
 from jax.test_util import check_grads
 
-import jax_linop
+import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
 
 # (x,y) -> (xy, y**2)
 def nonlin(out, args, kwargs_dump):
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     x, y = args
     out[0][()] = x * y
     out[1][()] = y * y
@@ -21,7 +21,7 @@ def nonlin(out, args, kwargs_dump):
 
 # (x,y,dx,dy) -> (ydx + xdy, 2 * y * dy)
 def nonlin_deriv(out, args, kwargs_dump):
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     x, y, dx, dy = args
     out[0][()] = y * dx + x * dy
     out[1][()] = 2 * y * dy
@@ -29,7 +29,7 @@ def nonlin_deriv(out, args, kwargs_dump):
 
 # (x, y, da, db) -> (yda, xda + 2ydb)
 def nonlin_deriv_T(out, args, kwargs_dump):
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     x, y, da, db = args
     out[0][()] = y * da
     out[1][()] = x * da + 2 * y * db
@@ -46,7 +46,7 @@ def nonlin_abstract(*args, **kwargs):
 
 
 funcs_deriv = (nonlin_deriv, nonlin_deriv_T)
-nonlin_jax = jax_linop.get_nonlinear_call(
+nonlin_jax = jaxbind.get_nonlinear_call(
     nonlin,
     funcs_deriv,
     nonlin_abstract,
@@ -67,7 +67,7 @@ check_grads(
 ################################################## test non diff args
 # (x,y) -> (xy, y**2)
 def nonlin(out, args, kwargs_dump):
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     x, y = args
     out[0][()] = x * y
     out[1][()] = y * y
[…]

### 2024-03-11 demos/demo_scipy_fft_batching.py

@@ -4,7 +4,7 @@ from jax import random
 
 import scipy
 
-import jax_linop
+import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
@@ -29,7 +29,7 @@ jax.config.update("jax_enable_x64", True)
 # thus in our case this will be a tuple of length 1 containing the input array
 # for the scipy fft.
 # The third argument are potential keyword arguments given later to the JAX
-# primitive. Due to the internals of JAX and jax_linop these keyword arguments
+# primitive. Due to the internals of JAX and jaxbind these keyword arguments
 # are passed to the function in serialized from and need to be deserialized.
 # For functions supporting custom batching the keyword arguments also contain as
 # a list of the axis over which the function should be batched.
@@ -40,7 +40,7 @@ def fftn(out, args, kwargs_dump):
     (x,) = args
 
     # deserialize keyword arguments
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
 
     # extract keyword argument which can be given to the JAX primitive
     workers = kwargs.pop("workers", None)
@@ -58,14 +58,14 @@ def fftn(out, args, kwargs_dump):
     out[0][()] = scipy.fft.fftn(x, axes=axes, norm="forward", workers=workers)
 
 
-# Besides the application of the FFT jax_linop also requires the application of
+# Besides the application of the FFT jaxbind also requires the application of
 # the linear transposed function. The syntax is identical to fftn just computing
 # the linear transposed.
 
 
 def fftn_transposed(out, args, kwargs_dump):
     (x,) = args
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
 
     workers = kwargs.pop("workers", None)
     batch_axes = kwargs.pop("batch_axes", None)
@@ -134,8 +134,8 @@ def fftn_transposed_abstract_eval(*args, **kwargs):
 
 
 # Now we register our function as a custom JAX primitive using the interface for
-# linear functions of jax_linop. jax_linop returns the resulting JAX primitive.
-fftn_jax = jax_linop.get_linear_call(
+# linear functions of jaxbind. jaxbind returns the resulting JAX primitive.
+fftn_jax = jaxbind.get_linear_call(
     fftn,
     fftn_transposed,
     fftn_abstract_eval,

### 2024-03-11 demos/s2fft_demo.py

@@ -6,7 +6,7 @@ config.update("jax_enable_x64", True)
 import ducc0
 import numpy as np
 import s2fft
-import jax_linop
+import jaxbind
 from time import time
 from functools import partial
 
@@ -66,7 +66,7 @@ def ssht_operator(L, sampling, spin, reality, nthreads, nside=0):
     def ssht_T(out_, args, kwargs_dump):
         out, = out_
         inp, = args
-        state = jax_linop.load_kwargs(kwargs_dump)
+        state = jaxbind.load_kwargs(kwargs_dump)
         L, spin, nthreads, reality, geometry, mapshape, hpxparam = get_config(out, inp, state)
 
         if spin==0:
@@ -112,7 +112,7 @@ def ssht_operator(L, sampling, spin, reality, nthreads, nside=0):
     def ssht(out_, args, kwargs_dump):
         out, = out_
         inp, = args
-        state = jax_linop.load_kwargs(kwargs_dump)
+        state = jaxbind.load_kwargs(kwargs_dump)
         L, spin, nthreads, reality, geometry, mapshape, hpxparam = get_config(inp, out, state)
 
         tmp = ducc0.sht.experimental.flm2alm(inp, spin, real=reality)
@@ -206,7 +206,7 @@ def ssht_operator(L, sampling, spin, reality, nthreads, nside=0):
     sampling = str(sampling)
     if sampling!="mw" and sampling!="dh" and sampling!="mwss" and sampling!="healpix":
         raise ValueError("unsupported sampling type")
-    op = jax_linop.get_linear_call(
+    op = jaxbind.get_linear_call(
         ssht, ssht_T, ssht_abstract, ssht_abstract_T,
         func_can_batch=False)
     return partial(op,

### 2024-03-11 demos/test_reference.py

@@ -2,7 +2,7 @@
 import jax
 from jax import numpy as jnp
 
-import jax_linop
+import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
@@ -25,7 +25,7 @@ def lin_abstract(*args, **kwargs):
     return ((x.shape, x.dtype), (x.shape, x.dtype))
 
 
-lin_jax = jax_linop.get_linear_call(lin, lin_T, lin_abstract, lin_abstract)
+lin_jax = jaxbind.get_linear_call(lin, lin_T, lin_abstract, lin_abstract)
 inp = (4 + jnp.zeros((2, 2)), 1 + jnp.zeros((2, 2)))
 lin_jax = jax.jit(lin_jax, static_argnames=("axes",))
 

### 2024-03-11 jax_linop/__init__.py

@@ -1,4 +1,4 @@
 __version__ = "0.1.0"
 
-from .jax_linop import get_linear_call, get_nonlinear_call
+from .jaxbind import get_linear_call, get_nonlinear_call
 from .misc import load_kwargs

### 2024-03-11 jax_linop/jax_linop.py

@@ -16,9 +16,9 @@ from jax.interpreters import ad, batching, mlir
 from jax.interpreters.mlir import ir_constant as irc
 from jaxlib.hlo_helpers import custom_call
 
-import _jax_linop
+import _jaxbind
 
-for _name, _value in _jax_linop.registrations().items():
+for _name, _value in _bind.registrations().items():
     jax.lib.xla_client.register_custom_call_target(_name, _value, platform="cpu")
 
 
@@ -267,7 +267,7 @@ def _batch(args, in_axes, *, _func: FunctionType, **kwargs):
     return y, out_axes
 
 
-_prim = jax.core.Primitive("jax_linop_prim")
+_prim = jax.core.Primitive("jaxbind_prim")
 _prim.multiple_results = True
 _prim.def_impl(partial(jax.interpreters.xla.apply_primitive, _prim))
 _prim.def_abstract_eval(_exec_abstract)
@@ -306,7 +306,7 @@ def get_linear_call(
         The args tuple contains the input for the function. In kwargs_dump,
         potential keyword arguments are contained in serialized form. The
         keyword arguments can be deserialized via
-        `jax_linop.load_kwargs(kwargs_dump)`.
+        `jaxbind.load_kwargs(kwargs_dump)`.
     abstract, abstract_T : function respectively its transpose
         Computing the shape and dtype of the operator's output from shape and
         dtype of its input. Its signature must be `(*args, **kwargs)`. `args`
@@ -378,7 +378,7 @@ def get_nonlinear_call(
         The args tuple contains the input for the function. In kwargs_dump,
         potential keyword arguments are contained in serialized form. The
         keyword arguments can be deserialized via
-        `jax_linop.load_kwargs(kwargs_dump)`.
+        `jaxbind.load_kwargs(kwargs_dump)`.
     f_derivative: tuple of functions
         Tuple containing functions for evaluating jvp and vjp of f. The fist entry
         in the function should evaluate jvp, the second vjp. The signature of the

### 2024-03-11 setup.py

@@ -4,7 +4,7 @@ import codecs
 import pybind11
 from setuptools import find_packages, setup, Extension
 
-pkgname = 'jax_linop'
+pkgname = 'jaxbind'
 
 
 def read(rel_path):
@@ -34,16 +34,16 @@ if sys.platform == 'darwin':
 elif sys.platform == 'win32':
     extra_compile_args = ['/EHsc', '/std:c++17']
 
-extensions = [Extension("_jax_linop",
+extensions = [Extension("_jaxbind",
                         language='c++',
-                        sources=['src/_jax_linop.cc'],
-                        depends=['src/_jax_linop.cc', 'setup.py'],
+                        sources=['src/_jaxbind.cc'],
+                        depends=['src/_jaxbind.cc', 'setup.py'],
                         include_dirs=include_dirs,
                         extra_compile_args=extra_compile_args,
                         extra_link_args=python_module_link_args)]
 
 setup(name=pkgname,
       version=get_version(f"{pkgname}/__init__.py"),
-      packages=find_packages(include=["jax_linop"]),
+      packages=find_packages(include=["jaxbind"]),
       ext_modules=extensions
       )

### 2024-03-11 src/_jax_linop.cc

@@ -1,7 +1,7 @@
 /* SPDX-License-Identifier: BSD-3-Clause */
 
 /*
- *  Jax_linop is being developed at the Max-Planck-Institut fuer Astrophysik
+ *  JAXbind is being developed at the Max-Planck-Institut fuer Astrophysik
  */
 
 /*
@@ -112,7 +112,7 @@ pybind11::dict Registrations()
 
 }
 
-PYBIND11_MODULE(_jax_linop, m) {
+PYBIND11_MODULE(_jaxbind, m) {
   m.def("registrations", detail_pymodule_jax::Registrations);
 }
 

### 2024-03-11 test/test_batching.py

@@ -4,7 +4,7 @@ import pytest
 import scipy
 from jax.test_util import check_grads
 
-import jax_linop
+import jaxbind
 
 pmp = pytest.mark.parametrize
 
@@ -13,7 +13,7 @@ jax.config.update("jax_enable_x64", True)
 
 def f(out, args, kwargs_dump):
     x, y = args
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     workers = kwargs.pop("workers", None)
     batch_axes = kwargs.pop("batch_axes", None)
     axes0 = list(range(len(x.shape)))
@@ -27,7 +27,7 @@ def f(out, args, kwargs_dump):
 
 def f_T(out, args, kwargs_dump):
     x, y = args
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     workers = kwargs.pop("workers", None)
     batch_axes = kwargs.pop("batch_axes", None)
     axes0 = list(range(len(x.shape)))
@@ -71,8 +71,8 @@ def f_a_T(*args, **kwargs):
     return ((a.shape, a.dtype, out_ax[0]), (b.shape, b.dtype, out_ax[1]))
 
 
-f_jax = jax_linop.get_linear_call(f, f_T, f_a, f_a_T, func_can_batch=False)
-f_jax_can_batch = jax_linop.get_linear_call(f, f_T, f_a, f_a_T, func_can_batch=True)
+f_jax = jaxbind.get_linear_call(f, f_T, f_a, f_a_T, func_can_batch=False)
+f_jax_can_batch = jaxbind.get_linear_call(f, f_T, f_a, f_a_T, func_can_batch=True)
 
 
 rng = np.random.default_rng(42)

### 2024-03-11 test/test_lin.py

@@ -13,7 +13,7 @@ import pytest
 from jax.test_util import check_grads
 from numpy.testing import assert_allclose
 
-import jax_linop
+import jaxbind
 
 pmp = pytest.mark.parametrize
 
@@ -39,7 +39,7 @@ def complextype(dtype):
 
 
 def fhtfunc(out, args, kwargs_dump):
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     # This function must _not_ keep any reference to 'inp' or 'args'!
     # Also, it must not change 'args'.
     if have_ducc0:
@@ -55,7 +55,7 @@ def fhtfunc_abstract(*args, **kwargs):
 
 
 def c2cfunc(out, args, kwargs_dump):
-    kwargs = jax_linop.load_kwargs(kwargs_dump)
+    kwargs = jaxbind.load_kwargs(kwargs_dump)
     (x,) = args
     if have_ducc0:
         ducc0.fft.c2c(x, out=out[0], **kwargs)
@@ -91,7 +91,7 @@ def realalm2alm(alm, lmax, dtype, out=None):
 def sht2d_operator(lmax, mmax, ntheta, nphi, geometry, spin, nthreads):
     def sht2dfunc(out, args, kwargs_dump):
         inp = args[0]
-        state = jax_linop.load_kwargs(kwargs_dump)
+        state = jaxbind.load_kwargs(kwargs_dump)
         tmp = realalm2alm(inp, state["lmax"], complextype(inp.dtype))
         ducc0.sht.synthesis_2d(
             lmax=state["lmax"],
@@ -107,7 +107,7 @@ def sht2d_operator(lmax, mmax, ntheta, nphi, geometry, spin, nthreads):
 
     def sht2dfunc_T(out, args, kwargs_dump):
         inp = args[0]
-        state = jax_linop.load_kwargs(kwargs_dump)
+        state = jaxbind.load_kwargs(kwargs_dump)
         tmp = ducc0.sht.adjoint_synthesis_2d(
             lmax=state["lmax"],
             mmax=state["mmax"],
@@ -133,7 +133,7 @@ def sht2d_operator(lmax, mmax, ntheta, nphi, geometry, spin, nthreads):
         shape_out = (ncomp, nalm)
         return ((shape_out, args[0].dtype),)
 
-    func = jax_linop.get_linear_call(
[…]

### 2024-03-11 test/test_mlin.py

@@ -3,7 +3,7 @@ import numpy as np
 from jax import numpy as jnp
 from jax.test_util import check_grads
 
-import jax_linop
+import jaxbind
 
 from functools import partial
 
@@ -42,7 +42,7 @@ def test_mlin():
 
     func_T = (mlin_T1, mlin_T2)
     func_abstract_T = (mlin_abstract_T1, mlin_abstract_T2)
-    mlin_jax = jax_linop.get_linear_call(
+    mlin_jax = jaxbind.get_linear_call(
         mlin,
         func_T,
         mlin_abstract,
@@ -95,7 +95,7 @@ def test_mlin_fixed_arg():
 
     func_T = (None, mlin_T2)
     func_abstract_T = (None, mlin_abstract_T2)
-    mlin_jax = jax_linop.get_linear_call(
+    mlin_jax = jaxbind.get_linear_call(
         mlin,
         func_T,
         mlin_abstract,

### 2024-03-11 test/test_nonlin.py

@@ -6,7 +6,7 @@ import pytest
 from jax import random
 from jax.test_util import check_grads
 
-import jax_linop
+import jaxbind
 
 pmp = pytest.mark.parametrize
 
@@ -100,7 +100,7 @@ def test_derivatives(first_n_args_fixed, dtype, shape, seed, jit):
         )
         absr_T = f_abstract_vjp_fix_x
 
-    f_jax = jax_linop.get_nonlinear_call(
+    f_jax = jaxbind.get_nonlinear_call(
         f,
         funcs_deriv,
         f_abstract,
@@ -151,7 +151,7 @@ def test_vmap(in_axes, dtype, shape, seed, jit):
     )
     absr_T = f_abstract
 
-    f_jax = jax_linop.get_nonlinear_call(
+    f_jax = jaxbind.get_nonlinear_call(
         f,
         funcs_deriv,
         f_abstract,

### 2024-03-11 jaxbind/jaxbind.py
jax_linop to jaxbind fixup

@@ -18,7 +18,7 @@ from jaxlib.hlo_helpers import custom_call
 
 import _jaxbind
 
-for _name, _value in _bind.registrations().items():
+for _name, _value in _jaxbind.registrations().items():
     jax.lib.xla_client.register_custom_call_target(_name, _value, platform="cpu")
 
 

### 2024-03-11 demos/demo_scipy_fft_batching.py
jax_op to JAXbind

@@ -8,21 +8,21 @@ import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
-# This demo showcases the use of the interface of jax_op for  binding linear
+# This demo showcases the use of the interface of JAXbind for  binding linear
 # functions to jax. Specifically it wraps the scipy fft as a JAX primitive.
 
 # All JAX primitives can batched over input axes with 'jax.vmap'. JAX primitives
-# registered via jax_op are no exception from that. By default jax_op internally
+# registered via JAXbind are no exception from that. By default JAXbind internally
 # translates the mapping operation into a sequential application of the function
 # along the batching axis. Nevertheless, some custom function such as the scipy
 # FFT natively support mapping over input axis giving significant speedups
-# compared to a sequential computation. The jax_op interface allows to makes use
+# compared to a sequential computation. The JAXbind interface allows to makes use
 # of custom batch, which will be demonstrated in this demo.
 
 
 # scipy.fft.fftn is the function we want to wrap as a JAX primitive. Therefore,
 # we wrap the scipy fft into a function 'fftn' having the signature required by
-# jax_op. jax_op requires the function to take 3 arguments.
+# JAXbind. JAXbind requires the function to take 3 arguments.
 # The fist argument 'out' is a tuple into which the result is written, thus in
 # this case an array containing the output of scipy.fft.fftn.
 # The second argument is also a tuple and contains the input for the function,

### 2024-03-11 jaxbind/jaxbind.py
add __all__

@@ -16,6 +16,9 @@ from jax.interpreters import ad, batching, mlir
 from jax.interpreters.mlir import ir_constant as irc
 from jaxlib.hlo_helpers import custom_call
 
+
+__all__ = ['get_linear_call', 'get_nonlinear_call']
+
 import _jaxbind
 
 for _name, _value in _jaxbind.registrations().items():
@@ -179,8 +182,9 @@ def _jvp(args, tangents, *, _func: FunctionType, **kwargs):
     return (res, tans)
 
 
-# NOTE: for whatever reason JAX will pass each arg separately to _transpose and not
-# as a tuple as for _jvp. Thus we need *args since we don't know the number of arguments.
+# NOTE: for whatever reason JAX will pass each arg separately to _transpose
+# and not as a tuple as for _jvp. Thus we need *args since we don't know
+# the number of arguments.
 def _transpose(cotangents, *args, _func: FunctionType, **kwargs):
     assert isinstance(_func, (LinearFunction, MultiLinearFunction))
     if _func.T is None:

### 2024-03-11 demos/demo_nonlin.py
BSD-2-Clause License

@@ -1,3 +1,8 @@
+# SPDX-License-Identifier: BSD-2-Clause
+# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
+
+# Copyright(C) 2024 Max-Planck-Society
+
 # %%
 from functools import partial
 

### 2024-03-11 demos/demo_scipy_fft_batching.py

@@ -1,3 +1,8 @@
+# SPDX-License-Identifier: BSD-2-Clause
+# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
+
+# Copyright(C) 2024 Max-Planck-Society
+
 import jax
 import jax.numpy as jnp
 from jax import random

### 2024-03-11 demos/s2fft_demo.py

@@ -1,3 +1,8 @@
+# SPDX-License-Identifier: BSD-2-Clause
+# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
+
+# Copyright(C) 2024 Max-Planck-Society
+
 import jax
 
 from jax import config

### 2024-03-11 demos/test_reference.py

@@ -1,3 +1,8 @@
+# SPDX-License-Identifier: BSD-2-Clause
+# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
+
+# Copyright(C) 2024 Max-Planck-Society
+
 # %%
 import jax
 from jax import numpy as jnp

### 2024-03-11 jaxbind/custom_map.py

@@ -1,4 +1,4 @@
-# Copyright(C) 2022-2023 Gordian Edenhofer
+# Copyright(C) 2022-2024 Gordian Edenhofer
 # SPDX-License-Identifier: GPL-2.0+ OR BSD-2-Clause
 
 from functools import partial

### 2024-03-11 src/_jaxbind.cc

@@ -1,4 +1,4 @@
-/* SPDX-License-Identifier: BSD-3-Clause */
+/* SPDX-License-Identifier: BSD-2-Clause */
 
 /*
  *  JAXbind is being developed at the Max-Planck-Institut fuer Astrophysik

### 2024-03-11 test/test_batching.py

@@ -1,3 +1,8 @@
+# SPDX-License-Identifier: BSD-2-Clause
+# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
+
+# Copyright(C) 2024 Max-Planck-Society
+
 import jax
 import numpy as np
 import pytest

### 2024-03-11 test/test_lin.py

@@ -1,3 +1,8 @@
+# SPDX-License-Identifier: BSD-2-Clause
+# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
+
+# Copyright(C) 2024 Max-Planck-Society
+
 from functools import partial
 
 have_ducc0 = True

### 2024-03-11 test/test_mlin.py

@@ -1,3 +1,8 @@
+# SPDX-License-Identifier: BSD-2-Clause
+# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
+
+# Copyright(C) 2024 Max-Planck-Society
+
 import jax
 import numpy as np
 from jax import numpy as jnp

### 2024-03-11 test/test_nonlin.py

@@ -1,3 +1,8 @@
+# SPDX-License-Identifier: BSD-2-Clause
+# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
+
+# Copyright(C) 2024 Max-Planck-Society
+
 from functools import partial
 
 import jax

### 2024-03-11 demos/demo_scipy_fft_batching.py
demo_scipy_fft_batching.py: Tweak

@@ -3,6 +3,7 @@
 
 # Copyright(C) 2024 Max-Planck-Society
 
+# %%
 import jax
 import jax.numpy as jnp
 from jax import random
@@ -13,31 +14,33 @@ import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
-# This demo showcases the use of the interface of JAXbind for  binding linear
-# functions to jax. Specifically it wraps the scipy fft as a JAX primitive.
+# %% [markdown]
 
-# All JAX primitives can batched over input axes with 'jax.vmap'. JAX primitives
-# registered via JAXbind are no exception from that. By default JAXbind internally
-# translates the mapping operation into a sequential application of the function
-# along the batching axis. Nevertheless, some custom function such as the scipy
-# FFT natively support mapping over input axis giving significant speedups
-# compared to a sequential computation. The JAXbind interface allows to makes use
-# of custom batch, which will be demonstrated in this demo.
+# # Binding Scipy's FFT to JAX
 
+# This demo showcases the use of the interface of JAXbind for binding linear
+# functions to jax. Specifically, it wraps scipy's FFT as a JAX primitive.
 
-# scipy.fft.fftn is the function we want to wrap as a JAX primitive. Therefore,
-# we wrap the scipy fft into a function 'fftn' having the signature required by
-# JAXbind. JAXbind requires the function to take 3 arguments.
-# The fist argument 'out' is a tuple into which the result is written, thus in
-# this case an array containing the output of scipy.fft.fftn.
-# The second argument is also a tuple and contains the input for the function,
-# thus in our case this will be a tuple of length 1 containing the input array
-# for the scipy fft.
-# The third argument are potential keyword arguments given later to the JAX
-# primitive. Due to the internals of JAX and jaxbind these keyword arguments
-# are passed to the function in serialized from and need to be deserialized.
-# For functions supporting custom batching the keyword arguments also contain as
-# a list of the axis over which the function should be batched.
+# The code wraps the scipy.fft.fftn function as a JAX primitive using JAXbind.
+# It defines the fftn function with the required signature for JAXbind. The
+# fftn function takes three arguments: out, args, and kwargs_dump. It extracts
+# the input array from args, deserializes the keyword arguments from
+# kwargs_dump, and computes the FFT using scipy.fft.fftn. Similarly, it defines
+# the fftn_transposed function for the linear transposed operation. The code
[…]

### 2024-03-11 demos/demo_scipy_fft_batching.py
demo_scipy_fft_batching.py: Organize imports

@@ -6,9 +6,8 @@
 # %%
 import jax
 import jax.numpy as jnp
-from jax import random
-
 import scipy
+from jax import random
 
 import jaxbind
 

### 2024-03-11 demos/demo_nonlin.py
demo_nonlin.py: Rework

@@ -7,7 +7,6 @@
 from functools import partial
 
 import jax
-import numpy as np
 from jax import numpy as jnp
 from jax.test_util import check_grads
 
@@ -16,120 +15,88 @@ import jaxbind
 jax.config.update("jax_enable_x64", True)
 
 
-# (x,y) -> (xy, y**2)
+# %% [markdown]
+
+# # Binding non-linear functions to JAX with fixed arguments
+
+# Bind a nonlinear function to JAX using JAXbind. Some input parts are fixed
+# (nondifferentiable) to showcase JAXbind's handling of fixed arguments.
+
+# The script begins by defining the nonlinear function using the JAXbind
+# compatible signature of out, args, and kwargs_dump. The derivative
+# function, which is registered with JAX, has the same signature as the
+# nonlinear function but includes an additional variable for the tangent of
+# the variable being differentiated. Additionally, we define the transpose of
+# the derivative function, as JAX may transpose it for operations like
+# retrieving the VJP.
+
+# %%
+
+
 def nonlin(out, args, kwargs_dump):
-    kwargs = jaxbind.load_kwargs(kwargs_dump)
+    # (x,y) -> (xy, y**2)
     x, y = args
     out[0][()] = x * y
     out[1][()] = y * y
 
 
-# (x,y,dx,dy) -> (ydx + xdy, 2 * y * dy)
 def nonlin_deriv(out, args, kwargs_dump):
-    kwargs = jaxbind.load_kwargs(kwargs_dump)
-    x, y, dx, dy = args
-    out[0][()] = y * dx + x * dy
-    out[1][()] = 2 * y * dy
-
-
-# (x, y, da, db) -> (yda, xda + 2ydb)
-def nonlin_deriv_T(out, args, kwargs_dump):
-    kwargs = jaxbind.load_kwargs(kwargs_dump)
-    x, y, da, db = args
-    out[0][()] = y * da
-    out[1][()] = x * da + 2 * y * db
-
-
-def nonlin_abstract(*args, **kwargs):
-    # Returns `shape` and `dtype` of output as well as the added batch_axes of the `output``
-    out_axes = kwargs.pop("batch_axes", ())
[…]

### 2024-03-11 demos/demo_nonlin.py
demo_nonlin.py: Remove check_grads

@@ -8,7 +8,6 @@ from functools import partial
 
 import jax
 from jax import numpy as jnp
-from jax.test_util import check_grads
 
 import jaxbind
 
@@ -98,5 +97,3 @@ _ = jax.jvp(nonlin_jax_pt, (inp,), (inp,))
 _, nonlin_jax_pt_vjp = jax.vjp(nonlin_jax_pt, inp)
 nonlin_jax_pt_vjp = jax.jit(nonlin_jax_pt_vjp)
 _ = nonlin_jax_pt_vjp([inp, inp])
-
-check_grads(nonlin_jax_pt, (inp,), order=1, modes=("fwd", "rev"), eps=1e-3)

### 2024-03-11 demos/demo_nonlin.py
demo_nonlin.py: Remove debug print statements

@@ -63,7 +63,6 @@ def nonlin_deriv_T(out, args, kwargs_dump):
 def nonlin_abstract(*args, **kwargs):
     # Returns `shape` and `dtype` of output
     x, y, *_ = args
-    print(f"f {_=}")
     assert x.shape == x.shape and x.dtype is y.dtype
     return ((x.shape, x.dtype), (x.shape, x.dtype))
 
@@ -71,7 +70,6 @@ def nonlin_abstract(*args, **kwargs):
 def nonlin_abstract_T(*args, **kwargs):
     # Returns `shape` and `dtype` of output
     a, b, da, db = args
-    print(f"T {_=}")
     assert a.shape == b.shape == da.shape == db.shape
     assert a.dtype is b.dtype and a.dtype is da.dtype and a.dtype is db.dtype
     return ((a.shape, a.dtype),)

### 2024-03-11 demos/demo_mlin.py
demo_mlin: more comments

@@ -4,50 +4,91 @@
 # Copyright(C) 2024 Max-Planck-Society
 
 # %%
-from functools import partial
-
 import jax
-import numpy as np
-from jax import numpy as jnp
-from jax.test_util import check_grads
+import jax.numpy as jnp
+from jax import random
+
 
 import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
+# %% [markdown]
+
+# # Binding a multi-linear function to JAX
+
+# This demo showcases the use of JAXbind for binding multi-linear functions to
+# JAX. As an example we bind the python function mlin computing
+# (x,y) -> (x*y, x*y) to a JAX primitive. Note: multi-linear functions can also
+# be regarded als general non-linear functions. For the JAXbind interface for
+# non-linear functions see the 'demo_nonlin.py' and the docstring of the
+# jaxbind.get_nonlinear_cal'. Additional information for linear functions can
+# also be found in 'demo_scipy_fft.py'.
+
+
+
+# %%
+
 
 def mlin(out, args, kwargs_dump):
-    out[0][()] = args[0] * args[1]
-    out[1][()] = args[0] * args[1]
+    # extract the input from the input tuple
+    x,y = args[0], args[1]
 
+    # doe the computation and write result in the out tuple
+    out[0][()] = x*y
+    out[1][()] = x*y
 
+# %% [markdown]
+
+# Besides the application of the function ('mlin') itself, JAXbind required the
+# linear transposed of the partial derivatives of 'mlin'.
+# %%
+
+# linear transpose of the partial derivative of 'mlin' with respect to the fist
+# variable x.
 def mlin_T1(out, args, kwargs_dump):
-    out[0][()] = args[0] * args[1] + args[0] * args[2]
+    y, da, db = args[0], args[1], args[2]
+    out[0][()] = y * da + y * db
 
-
+# linear transpose of the partial derivative of 'mlin' with respect to the second
+# variable y.
 def mlin_T2(out, args, kwargs_dump):
-    out[0][()] = args[0] * args[1] + args[0] * args[2]
+    x, da, db = args[0], args[1], args[2]
[…]

### 2024-03-11 demos/demo_mlin_batch.py
demos: remove no longer needed demos

@@ -1,95 +1,0 @@
-# SPDX-License-Identifier: BSD-2-Clause
-# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
-
-# Copyright(C) 2024 Max-Planck-Society
-
-# %%
-from functools import partial
-
-import jax
-import numpy as np
-from jax import numpy as jnp
-from jax.test_util import check_grads
-
-import jaxbind
-
-jax.config.update("jax_enable_x64", True)
-
-
-def mlin_call(x, y):
-    return x * y, x * y
-
-
-def mlin(out, args, kwargs_dump):
-    kwargs = jaxbind.load_kwargs(kwargs_dump)
-    batch_axes = kwargs.pop("batch_axes", None)
-    call = mlin_call
-    if batch_axes is not None and batch_axes != ((),) * len(args):
-        # Allow vmapping exactly once
-        assert all(len(ba) in (0, 1) for ba in batch_axes)
-        in_axes = tuple((ba[0] if len(ba) == 1 else None) for ba in batch_axes)
-        call = jax.vmap(mlin_call, in_axes=in_axes)
-    o = call(*args)
-    out[0][()] = o[0]
-    out[1][()] = o[1]
-
-
-def mlin_T1(out, args, kwargs_dump):
-    out[0][()] = args[0] * args[1] + args[0] * args[2]
-
-
-def mlin_T2(out, args, kwargs_dump):
-    out[0][()] = args[0] * args[1] + args[0] * args[2]
-
-
-def mlin_abstract(*args, **kwargs):
-    # Returns `shape` and `dtype` of output as well as the added batch_axes of the `output``
-    batch_axes = kwargs.pop("batch_axes", None)
-    call = mlin_call
-    if batch_axes is not None and batch_axes != ((),) * len(args):
-        # Allow vmapping exactly once
-        assert all(len(ba) in (0, 1) for ba in batch_axes)
-        in_axes = tuple((ba[0] if len(ba) == 1 else None) for ba in batch_axes)
-        call = jax.vmap(mlin_call, in_axes=in_axes)
-    out_axes = [0 for ba in batch_axes]
-    out = jax.eval_shape(call, *args)
-    return tuple((o.shape, o.dtype, oa) for o, oa in zip(out, out_axes))
-
-
[…]

### 2024-03-11 demos/s2fft_demo.py

@@ -1,269 +1,0 @@
-# SPDX-License-Identifier: BSD-2-Clause
-# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
-
-# Copyright(C) 2024 Max-Planck-Society
-
-import jax
-
-from jax import config
-config.update("jax_enable_x64", True)
-
-import ducc0
-import numpy as np
-import s2fft
-import jaxbind
-from time import time
-from functools import partial
-
-r2cdict = {np.dtype(np.float32) : np.dtype(np.complex64),
-           np.dtype(np.float64) : np.dtype(np.complex128)}
-
-c2rdict = {np.dtype(np.complex64) : np.dtype(np.float32),
-           np.dtype(np.complex128) : np.dtype(np.float64)}
-
-def isrealtype(dtype):
-    return np.dtype(dtype) in r2cdict
-
-def iscomplextype(dtype):
-    return np.dtype(dtype) in c2rdict
-
-def isvalidtype(dtype):
-    return isrealtype(dtype) or iscomplextype(dtype)
-
-def realtype(dtype):
-    return c2rdict[np.dtype(dtype)]
-
-def complextype(dtype):
-    return r2cdict[np.dtype(dtype)]
-
-def ssht_operator(L, sampling, spin, reality, nthreads, nside=0):
-    def get_config(alm, map, state):
-        if not iscomplextype(alm.dtype):
-            raise RuntimeError("invalid alm data type")
-        if not isvalidtype(map.dtype):
-            raise RuntimeError("invalid map data type")
-        reality = state["reality"]
-        L = state["L"]
-        spin = state["spin"]
-        nthreads = state["nthreads"]
-        sampling = state["sampling"]
-        hpxparam={}
-        if sampling == "mw":
-            geometry = "MW"
-            mapshape = (L, 2*L-1)
-        elif sampling == "mwss":
-            geometry = "CC"
-            mapshape = (L+1, 2*L)
-        elif sampling == "dh":
-            geometry = "F1"
-            mapshape = (2*L, 2*L-1)
-        elif sampling == "healpix":
-            geometry = "healpix"
-            mapshape = (12*state["nside"]**2,)
-            base = ducc0.healpix.Healpix_Base(state["nside"], "RING")
-            hpxparam = base.sht_info()
-        if map.shape!=mapshape:
-            raise RuntimeError("bad map shape")
-        if alm.shape!=(L,2*L-1):
-            raise RuntimeError("bad alm shape")
[…]

### 2024-03-12 test/test_reference.py
test_reference: turn into test

@@ -11,37 +11,37 @@ import jaxbind
 
 jax.config.update("jax_enable_x64", True)
 
-
-def lin(out, args, kwargs_dump):
-    x, y = args
-    out[0][()] = x + y
-    out[1][()] = x + y
-
-
-def lin_T(out, args, kwargs_dump):
-    a, b = args
-    out[0][()] = a + b
-    out[1][()] = a + b
+def test_reference():
+    def lin(out, args, kwargs_dump):
+        x, y = args
+        out[0][()] = x + y
+        out[1][()] = x + y
 
 
-def lin_abstract(*args, **kwargs):
-    x, y = args
-    assert x.shape == y.shape
-    return ((x.shape, x.dtype), (x.shape, x.dtype))
+    def lin_T(out, args, kwargs_dump):
+        a, b = args
+        out[0][()] = a + b
+        out[1][()] = a + b
 
 
-lin_jax = jaxbind.get_linear_call(lin, lin_T, lin_abstract, lin_abstract)
-inp = (4 + jnp.zeros((2, 2)), 1 + jnp.zeros((2, 2)))
-lin_jax = jax.jit(lin_jax, static_argnames=("axes",))
-
-del lin, lin_abstract, lin_T
-
-
-def lin(*a):
-    raise ValueError()
+    def lin_abstract(*args, **kwargs):
+        x, y = args
+        assert x.shape == y.shape
+        return ((x.shape, x.dtype), (x.shape, x.dtype))
 
 
-_ = {"a": jnp.arange(10)}
+    lin_jax = jaxbind.get_linear_call(lin, lin_T, lin_abstract, lin_abstract)
+    inp = (4 + jnp.zeros((2, 2)), 1 + jnp.zeros((2, 2)))
+    lin_jax = jax.jit(lin_jax, static_argnames=("axes",))
 
-lin_jax(*inp, axes=(3, 4))
-jax.linear_transpose(lin_jax, *inp)(list(inp))
+    del lin, lin_abstract, lin_T
+
+
+    def lin(*a):
+        raise ValueError()
+
+
+    _ = {"a": jnp.arange(10)}
+
+    lin_jax(*inp, axes=(3, 4))
+    jax.linear_transpose(lin_jax, *inp)(list(inp))

### 2024-03-12 test/test_batching.py
test_batching: test derivatives of custom batching

@@ -102,7 +102,7 @@ def test_vmap(bt_a1, bt_a2, bt2_a1, bt2_a2, o_a1, o_a2):
     np.testing.assert_allclose(rj, rb)
     assert rj[0].shape == rb[0].shape and rj[1].shape == rb[1].shape
     # check_grads(vj, a, order=2, modes=["fwd", "rev"], eps=1.0)
-    # check_grads(vb, a, order=2, modes=["fwd", "rev"], eps=1.0)
+    check_grads(vb, a, order=2, modes=["fwd", "rev"], eps=1.0)
 
     vvj = jax.vmap(vj, in_axes=(bt2_a1, bt2_a2), out_axes=[o_a1, o_a2])
     vvb = jax.vmap(vb, in_axes=(bt2_a1, bt2_a2), out_axes=[o_a1, o_a2])

### 2024-03-12 test/test_batching.py
test_batching: test batch axis reordering

@@ -84,14 +84,14 @@ rng = np.random.default_rng(42)
 a1 = rng.random((5, 5, 5)) - 0.5 + 1j * (rng.random((5, 5, 5)) - 0.5)
 a2 = rng.random((5, 5, 5)) - 0.5 + 1j * (rng.random((5, 5, 5)) - 0.5)
 a = (a1, a2)
-av1 = rng.random((5, 5, 5, 3, 4)) - 0.5 + 1j * (rng.random((5, 5, 5, 3, 4)) - 0.5)
-av2 = rng.random((5, 5, 5, 4, 3)) - 0.5 + 1j * (rng.random((5, 5, 5, 4, 3)) - 0.5)
+av1 = rng.random((5, 5, 5, 5, 5)) - 0.5 + 1j * (rng.random((5, 5, 5, 5, 5)) - 0.5)
+av2 = rng.random((5, 5, 5, 5, 5)) - 0.5 + 1j * (rng.random((5, 5, 5, 5, 5)) - 0.5)
 av = (av1, av2)
 
 
 @pmp("bt_a1", (0, 1, 2))
 @pmp("bt_a2", (0, 1, 2))
-@pmp("bt2_a1,bt2_a2", ((3, 4), (4, 3)))
+@pmp("bt2_a1,bt2_a2", ((3, 4), (4, 3), (0,0)))
 @pmp("o_a1", (0, 1))
 @pmp("o_a2", (0, 1))
 def test_vmap(bt_a1, bt_a2, bt2_a1, bt2_a2, o_a1, o_a2):

### 2024-03-12 demos/demo_mlin.py
demo_mlin, test_batching: format

@@ -26,17 +26,17 @@ jax.config.update("jax_enable_x64", True)
 # also be found in 'demo_scipy_fft.py'.
 
 
-
 # %%
 
 
 def mlin(out, args, kwargs_dump):
     # extract the input from the input tuple
-    x,y = args[0], args[1]
+    x, y = args[0], args[1]
 
-    # doe the computation and write result in the out tuple
-    out[0][()] = x*y
-    out[1][()] = x*y
+    # do the computation and write result in the out tuple
+    out[0][()] = x * y
+    out[1][()] = x * y
+
 
 # %% [markdown]
 
@@ -44,17 +44,20 @@ def mlin(out, args, kwargs_dump):
 # linear transposed of the partial derivatives of 'mlin'.
 # %%
 
+
 # linear transpose of the partial derivative of 'mlin' with respect to the fist
 # variable x.
 def mlin_T1(out, args, kwargs_dump):
     y, da, db = args[0], args[1], args[2]
     out[0][()] = y * da + y * db
 
+
 # linear transpose of the partial derivative of 'mlin' with respect to the second
 # variable y.
 def mlin_T2(out, args, kwargs_dump):
     x, da, db = args[0], args[1], args[2]
     out[0][()] = x * da + x * db
+
 
 # %% [markdown]
 
@@ -64,6 +67,7 @@ def mlin_T2(out, args, kwargs_dump):
 # mlin_T1, and mlin_T2. The abstract evaluations functions return for each
 # output argument a tuple containing the shape and dtype of this output. More
 # details are in the 'demo_scipy_fft.py'
+
 
 # %%
 def mlin_abstract(*args, **kwargs):
@@ -82,6 +86,7 @@ def mlin_abstract_T1(*args, **kwargs):
 def mlin_abstract_T2(*args, **kwargs):
     assert args[0].shape == args[1].shape
     return ((args[0].shape, args[0].dtype),)
+
 
 # %% [markdown]
 
@@ -116,4 +121,3 @@ res_jit = mlin_jit(*inp)
 
 # compute the jvp
 res_jvp = jax.jvp(mlin_jit, inp, inp)
-

### 2024-03-12 test/test_batching.py

@@ -91,7 +91,7 @@ av = (av1, av2)
 
 @pmp("bt_a1", (0, 1, 2))
 @pmp("bt_a2", (0, 1, 2))
-@pmp("bt2_a1,bt2_a2", ((3, 4), (4, 3), (0,0)))
+@pmp("bt2_a1,bt2_a2", ((3, 4), (4, 3), (0, 0)))
 @pmp("o_a1", (0, 1))
 @pmp("o_a2", (0, 1))
 def test_vmap(bt_a1, bt_a2, bt2_a1, bt2_a2, o_a1, o_a2):

### 2024-03-12 demos/demo_lin.py
demo_lin: delete no longer needed demo

@@ -1,100 +1,0 @@
-# SPDX-License-Identifier: BSD-2-Clause
-# Authors: Martin Reinecke, Jakob Roth, Gordian Edenhofer
-
-# Copyright(C) 2024 Max-Planck-Society
-
-# %%
-from functools import partial
-
-import jax
-from jax import numpy as jnp
-from jax.test_util import check_grads
-
-import jaxbind
-
-jax.config.update("jax_enable_x64", True)
-
-
-def lin(out, args, kwargs_dump):
-    # NOTE, this might look inefficient but for most practical problems it
-    # really is not:
-    # ```
-    # kwargs_dump = np.frombuffer(pickle.dumps({"asdas": (1, 2, 3,)}), dtype=np.uint8)
-    # %timeit pickle.loads(np.ndarray.tobytes(kwargs_dump))
-    # # 582 ns ± 1.12 ns per loop (mean ± std. dev. of 7 runs, 1,000,000 loops each)
-    # ```
-    kwargs = jaxbind.load_kwargs(kwargs_dump)
-    x, y = args
-    out[0][()] = x + y
-    out[1][()] = x + y
-
-
-def lin_T(out, args, kwargs_dump):
-    a, b = args
-    out[0][()] = a + b
-    out[1][()] = a + b
-
-
-def lin_abstract(*args, **kwargs):
-    # Returns `shape` and `dtype` of output as well as the added batch_axes of the `output``
-    x, y = args
-    assert x.shape == y.shape
-    return ((x.shape, x.dtype), (x.shape, x.dtype))
-
-
-def lin_abstract_T(*args, **kwargs):
-    # Returns `shape` and `dtype` of output as well as the added batch_axes of the `output``
-    a, b = args
-    assert a.shape == b.shape
-    return ((a.shape, a.dtype), (a.shape, a.dtype))
-
-
-lin_jax = jaxbind.get_linear_call(
-    lin, lin_T, lin_abstract, lin_abstract_T, first_n_args_fixed=0,
-)
-inp = (4 + jnp.zeros((2, 2)), 1 + jnp.zeros((2, 2)))
-lin_jax(*inp, axes=(3, 4))
-
-check_grads(partial(lin_jax, axes=(3, 4)), inp, order=2, modes=["fwd", "rev"], eps=1.0)
-
-
-##################################### check fixing args #######################
-def lin(out, args, kwargs_dump):
-    x, y = args
[…]

### 2024-03-12 test/test_batching.py
test_batching: test different ax length

@@ -88,6 +88,9 @@ av1 = rng.random((5, 5, 5, 5, 5)) - 0.5 + 1j * (rng.random((5, 5, 5, 5, 5)) - 0.
 av2 = rng.random((5, 5, 5, 5, 5)) - 0.5 + 1j * (rng.random((5, 5, 5, 5, 5)) - 0.5)
 av = (av1, av2)
 
+av1_diff_ax_len = rng.random((5, 5, 5, 3, 4)) - 0.5 + 1j * (rng.random((5, 5, 5, 3, 4)) - 0.5)
+av2_diff_ax_len = rng.random((5, 5, 5, 4, 3)) - 0.5 + 1j * (rng.random((5, 5, 5, 4, 3)) - 0.5)
+av_diff_ax_len = (av1_diff_ax_len, av2_diff_ax_len)
 
 @pmp("bt_a1", (0, 1, 2))
 @pmp("bt_a2", (0, 1, 2))
@@ -113,6 +116,15 @@ def test_vmap(bt_a1, bt_a2, bt2_a1, bt2_a2, o_a1, o_a2):
     assert rj[0].shape == rb[0].shape and rj[1].shape == rb[1].shape
     # check_grads(vvj, av, order=2, modes=["fwd", "rev"], eps=1.0)
     # check_grads(vvb, av, order=2, modes=["fwd", "rev"], eps=1.0)
+
+    if not (bt2_a1,bt2_a2) ==(0, 0):
+        vvj = jax.vmap(vj, in_axes=(bt2_a1, bt2_a2), out_axes=[o_a1, o_a2])
+        vvb = jax.vmap(vb, in_axes=(bt2_a1, bt2_a2), out_axes=[o_a1, o_a2])
+        rb = vvb(*av_diff_ax_len)
+        rj = vvj(*av_diff_ax_len)
+        np.testing.assert_allclose(rj[0], rb[0])
+        np.testing.assert_allclose(rj[1], rb[1])
+        assert rj[0].shape == rb[0].shape and rj[1].shape == rb[1].shape
 
 
 if __name__ == "__main__":

### 2024-03-12 setup.py
setup.py: Format

@@ -4,18 +4,18 @@ import codecs
 import pybind11
 from setuptools import find_packages, setup, Extension
 
-pkgname = 'jaxbind'
+pkgname = "jaxbind"
 
 
 def read(rel_path):
     here = os.path.abspath(os.path.dirname(__file__))
-    with codecs.open(os.path.join(here, rel_path), 'r') as fp:
+    with codecs.open(os.path.join(here, rel_path), "r") as fp:
         return fp.read()
 
 
 def get_version(rel_path):
     for line in read(rel_path).splitlines():
-        if line.startswith('__version__'):
+        if line.startswith("__version__"):
             delim = '"' if '"' in line else "'"
             return line.split(delim)[1]
     else:
@@ -24,26 +24,31 @@ def get_version(rel_path):
 
 include_dirs = [pybind11.get_include(True), pybind11.get_include(False)]
 
-extra_compile_args = ['-std=c++17', '-fvisibility=hidden']
+extra_compile_args = ["-std=c++17", "-fvisibility=hidden"]
 
 python_module_link_args = []
 
-if sys.platform == 'darwin':
-    extra_compile_args += ['-mmacosx-version-min=10.14']
-    python_module_link_args += ['-mmacosx-version-min=10.14']
-elif sys.platform == 'win32':
-    extra_compile_args = ['/EHsc', '/std:c++17']
+if sys.platform == "darwin":
+    extra_compile_args += ["-mmacosx-version-min=10.14"]
+    python_module_link_args += ["-mmacosx-version-min=10.14"]
+elif sys.platform == "win32":
+    extra_compile_args = ["/EHsc", "/std:c++17"]
 
-extensions = [Extension("_jaxbind",
-                        language='c++',
-                        sources=['src/_jaxbind.cc'],
-                        depends=['src/_jaxbind.cc', 'setup.py'],
-                        include_dirs=include_dirs,
-                        extra_compile_args=extra_compile_args,
-                        extra_link_args=python_module_link_args)]
+extensions = [
+    Extension(
+        "_jaxbind",
+        language="c++",
+        sources=["src/_jaxbind.cc"],
+        depends=["src/_jaxbind.cc", "setup.py"],
+        include_dirs=include_dirs,
+        extra_compile_args=extra_compile_args,
+        extra_link_args=python_module_link_args,
+    )
+]
 
-setup(name=pkgname,
-      version=get_version(f"{pkgname}/__init__.py"),
[…]

### 2024-03-12 setup.py
setup.py: Update pybind11.get_include call

@@ -22,7 +22,7 @@ def get_version(rel_path):
         raise RuntimeError("Unable to find version string.")
 
 
-include_dirs = [pybind11.get_include(True), pybind11.get_include(False)]
+include_dirs = [pybind11.get_include()]
 
 extra_compile_args = ["-std=c++17", "-fvisibility=hidden"]
 

### 2024-03-12 paper/paper.tex

@@
 \hypersetup{unicode=true,
-            pdftitle={jax\_op: Use any function in JAX},
+            pdftitle={JAXbind: Easy bindings to JAX},
             pdfborder={0 0 0},
@@
 
-\title{\texttt{jax\_op}: Use any function in JAX}
+\title{\texttt{JAXbind}: Easy bindings to JAX}
 
@@
 connecting custom code requires deep knowledge of JAX and its C++
-backend. The aim of \texttt{jax\_op} is to drastically lower the burden
+backend. The aim of \texttt{JAXbind} is to drastically lower the burden
 of connecting custom functions implemented in other programming
-languages to JAX. Specifically, \texttt{jax\_op} provides an easy-to-use
+languages to JAX. Specifically, \texttt{JAXbind} provides an easy-to-use
 Python interface for defining custom, so-called JAX primitives
@@
 
-In this paper, we present \texttt{jax\_op}, a package for bridging any
+In this paper, we present \texttt{JAXbind}, a package for bridging any
 function to JAX without in-depth knowledge of JAX's transformation
@@
 
-We believe \texttt{jax\_op} to be highly useful in scientific computing.
+We believe \texttt{JAXbind} to be highly useful in scientific computing.
 We intend to use this package to connect the Hartley transform and the
@@
 to connect the nonlinear function \(f(x_1,x_2) = x_1x_2^2\) to JAX. The
-\texttt{jax\_op} package expects the Python function for \(f\) to take
+\texttt{JAXbind} package expects the Python function for \(f\) to take
 three positional arguments. The first argument, \texttt{out}, is a
@@
 \begin{Highlighting}[]
-\ImportTok{import}\NormalTok{ jax\_linop }\CommentTok{\#FIXME}
+\ImportTok{import}\NormalTok{ jaxbind }
 
 \KeywordTok{def}\NormalTok{ f(out, args, kwargs\_dump):}
-\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jax\_linop.load\_kwargs(kwargs\_dump)}
+\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jaxbind.load\_kwargs(kwargs\_dump)}
 \NormalTok{    x1, x2 }\OperatorTok{=}\NormalTok{ args}
@@
 \texttt{jvp}. Instead, an implementation of the pushforward has to be
-provided, which \texttt{jax\_op} will register as the \texttt{jvp} of
+provided, which \texttt{JAXbind} will register as the \texttt{jvp} of
 the JAX primitive of \(f\). For our example, this
@@
 \KeywordTok{def}\NormalTok{ f\_jvp(out, args, kwargs\_dump):}
-\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jax\_linop.load\_kwargs(kwargs\_dump)}
+\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jaxbind.load\_kwargs(kwargs\_dump)}
 \NormalTok{    x1, x2, dx1, dx2 }\OperatorTok{=}\NormalTok{ args}
@@
 \KeywordTok{def}\NormalTok{ f\_vjp(out, args, kwargs\_dump):}
-\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jax\_linop.load\_kwargs(kwargs\_dump)}
+\NormalTok{    kwargs }\OperatorTok{=}\NormalTok{ jaxbind.load\_kwargs(kwargs\_dump)}
 \NormalTok{    x1, x2, dy }\OperatorTok{=}\NormalTok{ args}
@@
 We have now defined all ingredients necessary to register a JAX
-primitive for our function \(f\) using the \texttt{jax\_op} package.
+primitive for our function \(f\) using the \texttt{JAXbind} package.
 
@@
 \begin{Highlighting}[]
-\NormalTok{f\_jax }\OperatorTok{=}\NormalTok{ jax\_linop.get\_nonlinear\_call(}
+\NormalTok{f\_jax }\OperatorTok{=}\NormalTok{ jaxbind.get\_nonlinear\_call(}
 \NormalTok{    f, (f\_jvp, f\_vjp), f\_abstract, f\_abstract\_T}
@@
 
-\texttt{f\_jax} is a JAX primitive registered via the \texttt{jax\_op}
+\texttt{f\_jax} is a JAX primitive registered via the \texttt{JAXbind}
 package supporting all JAX transformations. We can now compute the
@@
 implemented by the high-performance codes. Therefore, the current
-interface of \texttt{jax\_op} is, for simplicity, restricted to first
+interface of \texttt{JAXbind} is, for simplicity, restricted to first
 derivatives. In the future, the interface could be easily expanded if
@@
 derivatives can be expressed in terms of \(f\) and its transpose. To
-make use of these simplifications, \texttt{jax\_op} provides a special
+make use of these simplifications, \texttt{JAXbind} provides a special
 interface for linear functions, supporting higher order derivatives,
@@
 
-Currently, \texttt{jax\_op} only has CPU but no GPU support. With some
+Currently, \texttt{JAXbind} only has CPU but no GPU support. With some
 expertise on Python bindings for GPU kernels adding GPU support should

### 2024-03-12 setup.py
Unify installation

@@ -23,17 +23,13 @@ def get_version(rel_path):
 
 
 include_dirs = [pybind11.get_include()]
-
 extra_compile_args = ["-std=c++17", "-fvisibility=hidden"]
-
 python_module_link_args = []
-
 if sys.platform == "darwin":
     extra_compile_args += ["-mmacosx-version-min=10.14"]
     python_module_link_args += ["-mmacosx-version-min=10.14"]
 elif sys.platform == "win32":
     extra_compile_args = ["/EHsc", "/std:c++17"]
-
 extensions = [
     Extension(
         "_jaxbind",

### 2024-03-12 paper/paper.tex
paper.tex: Sync
paper.tex: Fix affiliation order
paper: unify adresses
paper.tex: \"u for ü
paper: intermix -> mix
paper: describe arguments of abstract funcs

@@
           \author[1*]{Martin Reinecke}
-          \author[1, 2*]{Gordian Edenhofer}
+          \author[1, 2, 4*]{Gordian Edenhofer}
 
@@
 Garching, Germany}
+      \affil[4]{Department of Astrophysics, University of Vienna,
+      T\"urkenschanzstr. 17, A-1180 Vienna, Austria}
       \affil[*]{These authors contributed equally.}
@@
 accessible from within JAX. Rewriting these is often infeasible and/or
-inefficient. Ideally, we would like to intermix existing
+inefficient. Ideally, we would like to mix existing
 high-performance code with JAX code. However, connecting code to JAX
@@
 \begin{Highlighting}[]
-\ImportTok{import}\NormalTok{ jaxbind }
+\ImportTok{import}\NormalTok{ jaxbind}
 
@@
 does not need to be specified again.
+The abstract evaluation functions take normal positional and keyword arguments.
 

### 2024-03-12 paper/paper.tex
paper.tex: sync changes

@@
 
-JAX is widely used in machine learning and scientific computing.
-Scientific computing relies on existing high-performance code which we
-would ideally like to use in JAX. Reimplementing the existing code in
-JAX is often impractical and the existing interface in JAX for
-connecting custom code requires deep knowledge of JAX and its C++
-backend. The aim of \texttt{JAXbind} is to drastically lower the burden
-of connecting custom functions implemented in other programming
-languages to JAX. Specifically, \texttt{JAXbind} provides an easy-to-use
-Python interface for defining custom, so-called JAX primitives
-supporting any JAX transformations.
+JAX is widely used in machine learning and scientific computing, the latter of which often relies on existing high-performance code that we would ideally like to incorporate into JAX.
+Reimplementing the existing code in JAX is often impractical and the existing interface in JAX for binding custom code requires deep knowledge of JAX and its C++ backend.
+The goal of \texttt{JAXbind} is to drastically reduce the effort required to bind custom functions implemented in other programming languages to JAX.
+Specifically, \texttt{JAXbind} provides an easy-to-use Python interface for defining custom so-called JAX primitives that support arbitrary JAX transformations.
 
@@
 The use of JAX (Bradbury et al., 2018) is widespread in the natural
-sciences. JAX's powerful transformation system is of especially high
-interest. It enables retrieving arbitrary derivatives of functions,
-batch computations, and just-in-time code compilation for additional
-performance. Its transformation system relies on all constituents of the
-computation being written in JAX.
+sciences.
+Of particular interest is JAX's powerful transformation system.
+It enables to retrieve arbitrary derivatives of functions, batch computations, and just-in-time code for additional performance.
+Its transformation system relies on all components of the computation being written in JAX.
 
@@
 partial derivatives and their transpose functions as a JAX native call,
-a so-called primitive. Derivatives, compilation rules, and batching
-rules are automatically registered with JAX.
+a so-called primitive.
 
@@
 We intend to use this package to connect the Hartley transform and the
-spherical harmonic transform from ducc (Reinecke, 2024) to NIFTy
+spherical harmonic transform from DUCC (Reinecke, 2024) to the probabilistic programming package NIFTy
 (Edenhofer et al., 2024) as well as the radio interferometry response
-from ducc with \texttt{resolve} (Arras et al., 2024) for radio
+from DUCC with the radio astronomy package \texttt{resolve} (Arras et al., 2024) for radio
 astronomy. Furthermore, we intend to connect the non-uniform FFT from
-ducc with JAX for applications in strong-lensing astrophysics. We
+DUCC with JAX for applications in strong-lensing astrophysics. We
 envision many further applications within and outside of astrophysics.
@@
 three positional arguments. The first argument, \texttt{out}, is a
-\texttt{tuple} into which the function results are written. The second
+\texttt{tuple} into which the results are written. The second
 argument is also a \texttt{tuple} containing the input to the function,
-in our case, \(x_1\) and \(x_2\). Via \texttt{kwargs\_dump}, potential
-keyword arguments given to the later registered Jax primitive can be
-forwarded to \texttt{f} in serialized form.
+in our case, \(x_1\) and \(x_2\). Via \texttt{kwargs\_dump}, any
+keyword arguments given to the registered JAX primitive can be
+forwarded to $f$ in a serialized form.
 
@@
 To just-in-time compile the function, JAX needs to abstractly evaluate
-the code, i.e.~it needs to be able to know the shape and dtype of the
-output of the custom function given only the shape and dtype of the
+the code, i.e.~it needs to be able to infer the shape and dtype of the
+output of the function given only the shape and dtype of the
 input. We have to provide these abstract evaluation functions returning
-the output shape and dtype given an input shape and dtype for \texttt{f}
+the output shape and dtype given an input shape and dtype for $f$
 as well as for the \texttt{vjp} application. The output shape of the
-\texttt{jvp} is identical to the output shape of \texttt{f} itself and
+\texttt{jvp} is identical to the output shape of $f$ itself and
 does not need to be specified again.
@@
 derivatives are sufficient, and higher order derivatives are often not
-implemented by the high-performance codes. Therefore, the current
+implemented by high-performance codes. Therefore, the current
 interface of \texttt{JAXbind} is, for simplicity, restricted to first
@@
 
-Currently, \texttt{JAXbind} only has CPU but no GPU support. With some
-expertise on Python bindings for GPU kernels adding GPU support should
-be fairly simple. We especially want to highlight that the interfacing
-with the JAX automatic differentiation engine is identical for CPU and
-GPU.
+Currently, \texttt{JAXbind} only supports primitives that act on CPU memory.
+In the future, GPU support could be added analogous to the CPU backend.
+Note, the automatic differentiation in JAX is backend agnostic and would thus not require any additional bindings.
 

### 2024-03-12 paper/paper.tex
paper: tweak
paper.tex: Sync

@@
 \hypersetup{unicode=true,
-            pdftitle={JAXbind: Easy bindings to JAX},
+            pdftitle={JAXbind: Bind any function to JAX},
             pdfborder={0 0 0},
@@
 
-\title{\texttt{JAXbind}: Easy bindings to JAX}
+\title{\texttt{JAXbind}: Bind any function to JAX}
 
@@
 
-JAX is widely used in machine learning and scientific computing, the latter of which often relies on existing high-performance code that we would ideally like to incorporate into JAX.
-Reimplementing the existing code in JAX is often impractical and the existing interface in JAX for binding custom code requires deep knowledge of JAX and its C++ backend.
-The goal of \texttt{JAXbind} is to drastically reduce the effort required to bind custom functions implemented in other programming languages to JAX.
-Specifically, \texttt{JAXbind} provides an easy-to-use Python interface for defining custom so-called JAX primitives that support arbitrary JAX transformations.
+JAX is widely used in machine learning and scientific computing, the
+latter of which often relies on existing high-performance code that we
+would ideally like to incorporate into JAX. Reimplementing the existing
+code in JAX is often impractical and the existing interface in JAX for
+binding custom code requires deep knowledge of JAX and its C++ backend.
+The goal of \texttt{JAXbind} is to drastically reduce the effort
+required to bind custom functions implemented in other programming
+languages to JAX. Specifically, \texttt{JAXbind} provides an easy-to-use
+Python interface for defining custom so-called JAX primitives that
+support arbitrary JAX transformations.
 
@@
 The use of JAX (Bradbury et al., 2018) is widespread in the natural
-sciences.
-Of particular interest is JAX's powerful transformation system.
-It enables to retrieve arbitrary derivatives of functions, batch computations, and just-in-time code for additional performance.
-Its transformation system relies on all components of the computation being written in JAX.
+sciences. Of particular interest is JAX's powerful transformation
+system. It enables to retrieve arbitrary derivatives of functions, batch
+computations, and just-in-time compile code for additional performance.
+Its transformation system relies on all components of the computation
+being written in JAX.
 
@@
 accessible from within JAX. Rewriting these is often infeasible and/or
-inefficient. Ideally, we would like to mix existing
-high-performance code with JAX code. However, connecting code to JAX
-requires knowledge of the internals of JAX and its C++ backend.
+inefficient. Ideally, we would like to mix existing high-performance
+code with JAX code. However, connecting code to JAX requires knowledge
+of the internals of JAX and its C++ backend.
 
@@
 We intend to use this package to connect the Hartley transform and the
-spherical harmonic transform from DUCC (Reinecke, 2024) to the probabilistic programming package NIFTy
-(Edenhofer et al., 2024) as well as the radio interferometry response
-from DUCC with the radio astronomy package \texttt{resolve} (Arras et al., 2024) for radio
-astronomy. Furthermore, we intend to connect the non-uniform FFT from
-DUCC with JAX for applications in strong-lensing astrophysics. We
-envision many further applications within and outside of astrophysics.
+spherical harmonic transform from DUCC (Reinecke, 2024) to the
+probabilistic programming package NIFTy (Edenhofer et al., 2024) as well
+as the radio interferometry response from DUCC with the radio astronomy
+package \texttt{resolve} (Arras et al., 2024). Furthermore, we intend to
+connect the non-uniform FFT from DUCC with JAX for applications in
+strong-lensing astrophysics. We envision many further applications
+within and outside of astrophysics.
 
@@
 three positional arguments. The first argument, \texttt{out}, is a
-\texttt{tuple} into which the results are written. The second
-argument is also a \texttt{tuple} containing the input to the function,
-in our case, \(x_1\) and \(x_2\). Via \texttt{kwargs\_dump}, any
-keyword arguments given to the registered JAX primitive can be
-forwarded to $f$ in a serialized form.
+\texttt{tuple} into which the results are written. The second argument
+is also a \texttt{tuple} containing the input to the function, in our
+case, \(x_1\) and \(x_2\). Via \texttt{kwargs\_dump}, any keyword
+arguments given to the registered JAX primitive can be forwarded to
+\(f\) in a serialized form.
 
@@
 the code, i.e.~it needs to be able to infer the shape and dtype of the
-output of the function given only the shape and dtype of the
-input. We have to provide these abstract evaluation functions returning
-the output shape and dtype given an input shape and dtype for $f$
-as well as for the \texttt{vjp} application. The output shape of the
-\texttt{jvp} is identical to the output shape of $f$ itself and
-does not need to be specified again.
-The abstract evaluation functions take normal positional and keyword arguments.
+output of the function given only the shape and dtype of the input. We
+have to provide these abstract evaluation functions returning the output
+shape and dtype given an input shape and dtype for \(f\) as well as for
+the \texttt{vjp} application. The output shape of the \texttt{jvp} is
+identical to the output shape of \(f\) itself and does not need to be
+specified again. The abstract evaluation functions take normal
+positional and keyword arguments.
 
@@
 derivatives are sufficient, and higher order derivatives are often not
-implemented by high-performance codes. Therefore, the current
-interface of \texttt{JAXbind} is, for simplicity, restricted to first
-derivatives. In the future, the interface could be easily expanded if
-specific use cases require higher order derivatives.
+implemented by high-performance codes. Therefore, the current interface
+of \texttt{JAXbind} is, for simplicity, restricted to first derivatives.
+In the future, the interface could be easily expanded if specific use
+cases require higher order derivatives.
 
@@
 
-Currently, \texttt{JAXbind} only supports primitives that act on CPU memory.
-In the future, GPU support could be added analogous to the CPU backend.
-Note, the automatic differentiation in JAX is backend agnostic and would thus not require any additional bindings.
+Currently, \texttt{JAXbind} only supports primitives that act on CPU
+memory. In the future, GPU support could be added analogous to the CPU
+backend. The automatic differentiation in JAX is backend agnostic and
+would thus not require any additional bindings to work on the GPU.
 
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 

### 2024-03-13 jaxbind/__init__.py
Tag version 1.0.0

@@ -1,4 +1,4 @@
-__version__ = "0.1.0"
+__version__ = "1.0.0"
 
 from .jaxbind import get_linear_call, get_nonlinear_call
 from .misc import load_kwargs

### 2024-03-29 jaxbind/jaxbind.py
jaxbind: document JAXbind internal api

@@ -43,6 +43,27 @@ FunctionType = Union[LinearFunction, MultiLinearFunction, NonLinearFunction]
 
 
 def _exec_abstract(*args, _func: FunctionType, **kwargs):
+    """JAXbind internal: bridging the user abstract function to JAX
+
+
+    Parameters
+    ----------
+    *args : tuple of shaped arrays
+        Input for the 'abstract' function in _func
+    _func : FunctionType
+        namedtuple containing all arguments from 'get_linear_call' and
+        'get_nonlinear_call'.
+    **kwargs : dict
+        Additional keyword arguments forwarded to functions contained in _func.
+
+
+    Returns
+    -------
+    tuple : tuple of jax.core.ShapedArray
+        Indicates the shape and dtype of each output argument of the function
+        'f' in _func.
+
+    """
     if _func.can_batch:
         assert "batch_axes" not in kwargs
         kwargs["batch_axes"] = _func.batch_axes
@@ -63,6 +84,27 @@ _dtype_dict = {
 
 
 def _lowering(ctx, *args, _func: FunctionType, _platform="cpu", **kwargs):
+    """JAXbind internal: lowering the user function to JAX/XLA
+
+
+    Parameters
+    ----------
+    ctx : mlir.LoweringRuleContext
+    _func : FunctionType
+        namedtuple containing all arguments from 'get_linear_call' and
+        'get_nonlinear_call'.
+    _platform : string
+        Indicates the desired backend. For now only 'cpu' is supported.
+    **kwargs : dict
+        Additional keyword arguments forwarded to functions contained in _func.
+
+
+    Returns
+    -------
+    tuple : ir.OpResultList
+        Results of the function 'f' in _func as mlir object.
+
+    """
     operands = [irc(id(_func.f))]  # Pass the ID of the callable to C++
     operand_layouts = [()]
 
@@ -117,12 +159,38 @@ def _lowering(ctx, *args, _func: FunctionType, _platform="cpu", **kwargs):
 
 
 def _explicify_zeros(x):
+    """JAXbind internal: helper function instantiating array with zeros for any
+    ad.Zero in x.
+    """
     if isinstance(x, (tuple, list)):
         return [ad.instantiate_zeros(t) if isinstance(t, ad.Zero) else t for t in x]
     return ad.instantiate_zeros(x) if isinstance(x, ad.Zero) else x
 
 
 def _jvp(args, tangents, *, _func: FunctionType, **kwargs):
[…]

### 2024-04-01 jaxbind/jaxbind.py
jaxbind: tweak docstrings

@@ -43,7 +43,7 @@ FunctionType = Union[LinearFunction, MultiLinearFunction, NonLinearFunction]
 
 
 def _exec_abstract(*args, _func: FunctionType, **kwargs):
-    """JAXbind internal: bridging the user abstract function to JAX
+    """bridging the user abstract function to JAX
 
 
     Parameters
@@ -84,7 +84,7 @@ _dtype_dict = {
 
 
 def _lowering(ctx, *args, _func: FunctionType, _platform="cpu", **kwargs):
-    """JAXbind internal: lowering the user function to JAX/XLA
+    """lowering the user function to JAX/XLA
 
 
     Parameters
@@ -159,7 +159,7 @@ def _lowering(ctx, *args, _func: FunctionType, _platform="cpu", **kwargs):
 
 
 def _explicify_zeros(x):
-    """JAXbind internal: helper function instantiating array with zeros for any
+    """helper function instantiating array with zeros for any
     ad.Zero in x.
     """
     if isinstance(x, (tuple, list)):
@@ -168,15 +168,15 @@ def _explicify_zeros(x):
 
 
 def _jvp(args, tangents, *, _func: FunctionType, **kwargs):
-    """JAXbind internal: bridging the user derivative function to JAX
+    """bridging the user derivative function to JAX
 
 
     Parameters
     ----------
     args : tuple of arrays
-        Position at which the jacobian is computed.
+        Position at which the Jacobian is computed.
     tangents : tuple of arrays
-        Tangents to which the jacobian is applied.
+        Tangents to which the Jacobian is applied.
     _func : FunctionType
         namedtuple containing all arguments from 'get_linear_call' and
         'get_nonlinear_call'.
@@ -254,7 +254,7 @@ def _jvp(args, tangents, *, _func: FunctionType, **kwargs):
 # and not as a tuple as for _jvp. Thus we need *args since we don't know
 # the number of arguments.
 def _transpose(cotangents, *args, _func: FunctionType, **kwargs):
-    """JAXbind internal: bridging the user transpose function to JAX
+    """bridging the user transpose function to JAX
 
 
     Parameters
@@ -262,7 +262,7 @@ def _transpose(cotangents, *args, _func: FunctionType, **kwargs):
     cotangents : tuple of arrays
         Cotangents for the transpose/ vjp application.
     *args : tuple of arrays
-        Position at which the jacobian is computed.
+        Position at which the Jacobian is computed.
[…]

### 2024-04-01 docs/generate.sh
docs: generate docs with sphinx

@@ -1,0 +1,7 @@
+#!/usr/bin/sh
+
+set -e
+
+sphinx-apidoc -e -o docs/source/mod jaxbind
+# sphinx-build -b html docs/source/ docs/build/
+sphinx-build -b html docs/source/ .

### 2024-04-01 docs/source/conf.py

@@ -1,0 +1,56 @@
+import jaxbind
+
+needs_sphinx = '3.2.0'
+
+extensions = [
+    'sphinx.ext.napoleon',    # Support for NumPy and Google style docstrings
+    'sphinx.ext.imgmath',     # Render math as images
+    'sphinx.ext.viewcode',    # Add links to highlighted source code
+    'sphinx.ext.intersphinx', # Links to other sphinx docs (mostly numpy)
+    # 'myst_parser',            # Parse markdown
+]
+master_doc = 'index'
+
+myst_enable_extensions = [
+    "amsmath",
+    "dollarmath",
+    "strikethrough",
+    "tasklist",
+]
+
+intersphinx_mapping = {"numpy": ("https://numpy.org/doc/stable/", None),
+                       #"matplotlib": ('https://matplotlib.org/stable/', None),
+                       "ducc0": ("https://mtr.pages.mpcdf.de/ducc/", None),
+                       "scipy": ('https://docs.scipy.org/doc/scipy/reference/', None),
+                       }
+
+autodoc_default_options = {'special-members': '__init__'}
+
+napoleon_google_docstring = False
+napoleon_numpy_docstring = True
+napoleon_use_ivar = True
+napoleon_use_admonition_for_notes = True
+napoleon_use_admonition_for_examples = True
+napoleon_use_admonition_for_references = True
+napoleon_include_special_with_doc = True
+
+imgmath_embed = True
+
+project = u'JAXbind'
+copyright = u'2024, Max-Planck-Society'
+author = u'Jakob Roth, Martin Reinecke, Gordian Edenhofer'
+
+release = jaxbind.__version__
+version = release[:-2]
+
+language = "en"
+exclude_patterns = []
+add_module_names = False
+
+html_theme = "pydata_sphinx_theme"
+html_context = {"default_mode": "dark"}
+
+html_last_updated_fmt = '%b %d, %Y'
+
+exclude_patterns = ['mod/modules.rst']
+# exclude_patterns = ['mod/modules.rst', 'mod/nifty8.logger.rst']

### 2024-04-01 docs/generate.sh
docs/generate: write docs to _site/

@@ -4,4 +4,4 @@ set -e
 
 sphinx-apidoc -e -o docs/source/mod jaxbind
 # sphinx-build -b html docs/source/ docs/build/
-sphinx-build -b html docs/source/ .
+sphinx-build -b html docs/source/ _site/

### 2024-04-01 jaxbind/contrib/jaxducc0.py
jaxducc0: add docsting to wgridder

@@ -241,6 +241,36 @@ def get_wgridder(
     verbosity=0,
     **kwargs,
 ):
+    """Create a JAX primitive for the ducc0 wgridder
+
+    Parameters
+    ----------
+    pixsize_x, pixsize_y : float
+        Size of the pixels in radian.
+    npix_x, npix_y : int
+        Number of pixels.
+    epsilon : float
+        Sets the required accuracy of the wgridder evaluation.
+    nthreads : int
+        Sets the number of threads used for evaluation. Default 1.
+    flip_v : bool
+        Whether or not to flip the v coordinate of the visibilities. Default
+        `False`.
+    verbosity : int
+        Sets the verbosity of the wgridder. For 0 no print out, for >0 verbose
+        output. Default 0.
+    **kwargs : dict
+        Additional forwarded to ducc wgridder.
+
+    Returns
+    -------
+    op : Jax primitive evaluating the ducc wgridder.
+        The Jax primitive has the
+        signature `(uvw, freq, image)` with `uvw` being an (N, 3) array the uvw
+        coordinates of the visibilities in meter, `freq`  a 1D array with the
+        frequencies in Herz, and `image` a 2D arrays of shape `(npix_x, npix_y)`
+        with the sky brightness in Jansky per Steradian.
+    """
     wgridder = partial(
         _wgridder,
         pixsize_x=pixsize_x,

### 2024-04-08 src/_jaxbind.cc
_jaxbind.cc: Fix comment

@@ -99,7 +99,7 @@ void pycall(void *out_raw, void **in)
   size_t size_kwargs = *reinterpret_cast<uint64_t *>(in[idx++]);
   py::array py_kwargs (dtp_kwargs, size_kwargs, in[idx++], dummy);
 
-  // Execute the Python function implementing the linear operation
+  // Execute the Python function implementing the desired operation
   func(py_out, py_in, py_kwargs);
   }
 

### 2024-04-08 paper/paper.tex

@@
 
-To the best of our knowledge no other code currently exists for
-connecting generic functions to JAX. The package that comes the closest
-is Enzyme-JAX (EnzymeAD, 2024). Enzyme-JAX allows one to differentiate a
-C++ function with Enzyme (W. Moses \& Churavy, 2020; W. S. Moses et al.,
-2021, 2022) and connect it together with its derivative to JAX. However,
-it enforces the use of Enzyme for deriving derivatives and does not
-allow for connecting arbitrary code to JAX.
+The functionality of \texttt{JAXbind} extends the external callback
+functionality in JAX. Currently, \texttt{JAXbind}, akin to the external
+callback functions in JAX, briefly requires python's global interpreter
+lock (GIL) to call the user specified python function. In contrast to
+JAX's external callback functions, it allows for both a custom
+Jacobian-vector product and vector-Jacobian product. To the best of our
+knowledge no other code currently exists for connecting generic
+functions and both of their Jacobian products to JAX. The package that
+comes the closest is Enzyme-JAX (EnzymeAD, 2024). Enzyme-JAX allows one
+to differentiate a C++ function with Enzyme (W. Moses \& Churavy, 2020;
+W. S. Moses et al., 2021, 2022) and connect it together with its
+derivative to JAX. However, it enforces the use of Enzyme for deriving
+derivatives and does not allow for connecting arbitrary code to JAX.
 
@@
 
-\NormalTok{cotan }\OperatorTok{=}\NormalTok{ (jnp.full((}\DecValTok{4}\NormalTok{,}\DecValTok{3}\NormalTok{), }\FloatTok{6.}\NormalTok{),)}
+\NormalTok{cotan }\OperatorTok{=}\NormalTok{ [jnp.full((}\DecValTok{4}\NormalTok{,}\DecValTok{3}\NormalTok{), }\FloatTok{6.}\NormalTok{)]}
 \NormalTok{res, f\_vjp }\OperatorTok{=}\NormalTok{ jax.vjp(f\_jax, }\OperatorTok{*}\NormalTok{inp)}
@@
 Theory (NIFTy.re): A Library for Gaussian Processes and Variational
-Inference}}. \url{http://arxiv.org/abs/2402.16683}
+Inference}}. \url{https://doi.org/10.48550/arXiv.2402.16683}
 
@@
 on High Performance Computing, Networking, Storage and Analysis}.
-ISBN:~\href{https://worldcat.org/isbn/9784665454445}{9784665454445}
+\url{https://doi.org/10.1109/SC41404.2022.00065}
 

### 2024-06-25 paper/paper.tex
Sync tex

@@
 code in JAX is often impractical and the existing interface in JAX for
-binding custom code requires deep knowledge of JAX and its C++ backend.
-The goal of \texttt{JAXbind} is to drastically reduce the effort
-required to bind custom functions implemented in other programming
-languages to JAX. Specifically, \texttt{JAXbind} provides an easy-to-use
-Python interface for defining custom so-called JAX primitives that
-support arbitrary JAX transformations.
+binding custom code either limits the user to a single Jacobian product
+or requires deep knowledge of JAX and its C++ backend for general
+Jacobian products. With \texttt{JAXbind} we drastically reduce the
+effort required to bind custom functions implemented in other
+programming languages with full support for Jacobian-vector products and
+vector-Jacobian products to JAX. Specifically, \texttt{JAXbind} provides
+an easy-to-use Python interface for defining custom, so-called JAX
+primitives. Via \texttt{JAXbind}, any function callable from Python can
+be exposed as a JAX primitive. \texttt{JAXbind} allows a user to
+interface the JAX function transformation engine with custom derivatives
+and batching rules, enabling all JAX transformations for the custom
+primitive.
 
@@
 sciences. Of particular interest is JAX's powerful transformation
-system. It enables to retrieve arbitrary derivatives of functions, batch
-computations, and just-in-time compile code for additional performance.
-Its transformation system relies on all components of the computation
-being written in JAX.
+system. It enables a user to retrieve arbitrary derivatives of
+functions, batch computations, and just-in-time compile code for
+additional performance. Its transformation system requires that all
+components of the computation are written in JAX.
 
 A plethora of high-performance code is not written in JAX and thus not
-accessible from within JAX. Rewriting these is often infeasible and/or
-inefficient. Ideally, we would like to mix existing high-performance
-code with JAX code. However, connecting code to JAX requires knowledge
-of the internals of JAX and its C++ backend.
+accessible from within JAX. Rewriting these codes is often infeasible
+and/or inefficient. Ideally, we would like to mix existing
+high-performance code with JAX code. However, connecting code to JAX
+requires knowledge of the internals of JAX and its C++ backend.
 
@@
 system. The interface is accessible from Python without requiring any
-development in C++. The package is able to register any function, its
+development in C++. The package is able to register any function and its
 partial derivatives and their transpose functions as a JAX native call,
@@
 functionality in JAX. Currently, \texttt{JAXbind}, akin to the external
-callback functions in JAX, briefly requires python's global interpreter
-lock (GIL) to call the user specified python function. In contrast to
-JAX's external callback functions, it allows for both a custom
-Jacobian-vector product and vector-Jacobian product. To the best of our
-knowledge no other code currently exists for connecting generic
-functions and both of their Jacobian products to JAX. The package that
-comes the closest is Enzyme-JAX (EnzymeAD, 2024). Enzyme-JAX allows one
-to differentiate a C++ function with Enzyme (W. Moses \& Churavy, 2020;
-W. S. Moses et al., 2021, 2022) and connect it together with its
-derivative to JAX. However, it enforces the use of Enzyme for deriving
-derivatives and does not allow for connecting arbitrary code to JAX.
+callback functions in JAX, briefly requires Python's global interpreter
+lock (GIL) to call the user-specified Python function. In contrast to
+JAX's external callback functions, \texttt{JAXbind} allows for both a
+custom Jacobian-vector product and vector-Jacobian product. To the best
+of our knowledge no other code currently exists for easily binding
+generic functions and both of their Jacobian products to JAX, without
+the need for C++ or LLVM. The package that comes the closest is
+Enzyme-JAX (W. S. Moses \& Zinenko, 2024), which allows one to bind
+arbitrary LLVM/MLIR, including C++, with automatically-generated (W.
+Moses \& Churavy, 2020; W. S. Moses et al., 2021, 2022) or
+manually-defined derivatives to JAX.
+
+PyTorch (Ansel et al., 2024) and TensorFlow (Abadi et al., 2015) also
+provide interfaces for custom extensions. PyTorch has an extensively
+documented Python interface\footnote{\url{https://pytorch.org/docs/stable/notes/extending.html}}
+for wrapping custom Python functions as PyTorch functions. This
+interface connects the custom function to PyTorch's automatic
+differentiation engine, allowing for custom Jacobian and Jacobian
+transposed applications, similar to what is possible with JAXbind.
+Additionally, PyTorch allows a user to interface its C++ backend with
+custom C++ or CUDA extensions\footnote{\url{https://pytorch.org/tutorials/advanced/cpp_extension.html}}.
+JAXbind, in contrast, currently only supports functions executed on the
+CPU, although the JAX built-in C++ interface also allows for custom GPU
+kernels. TensorFlow includes a C++ interface\footnote{\url{https://www.tensorflow.org/guide/create_op}}
+for custom functions that can be executed on the CPU or GPU. Custom
+gradients can be added to these functions.
 
@@
 registered with JAX support automatic differentiation. In the following,
-we will outline which functions our package respectively JAX requires to
-enable automatic differentiation. For simplicity, we assume that we want
-to connect the nonlinear function \(f(x_1,x_2) = x_1x_2^2\) to JAX. The
+we will outline which functions our package requires to enable automatic
+differentiation via JAX. For simplicity, we assume that we want to
+connect the nonlinear function \(f(x_1,x_2) = x_1x_2^2\) to JAX. The
 \texttt{JAXbind} package expects the Python function for \(f\) to take
@@
 To just-in-time compile the function, JAX needs to abstractly evaluate
-the code, i.e.~it needs to be able to infer the shape and dtype of the
+the code, i.e., it needs to be able to infer the shape and dtype of the
 output of the function given only the shape and dtype of the input. We
@@
 differentiation becomes trivial. Specifically for a linear function
-\(f\), the pushforward respectively the \texttt{jvp} of \(f\) is
-identical to \(f\) itself and independent of the position at which it is
-computed. Expressed in formulas, \(\partial f(x)(dx) = f(dx)\) if \(f\)
-is linear in \(x\). Analogously, the pullback respectively the
-\texttt{vjp} becomes independent of the initial position and is given by
-the linear transpose of \(f\), thus
-\((\partial f(x))^{T}(dy) = f^T(dy)\). Also, all higher order
+\(f\), the pushforward or \texttt{jvp} of \(f\) is identical to \(f\)
+itself and independent of the position at which it is computed.
+Expressed in formulas, \(\partial f(x)(dx) = f(dx)\) if \(f\) is linear
+in \(x\). Analogously, the pullback or \texttt{vjp} becomes independent
+of the initial position and is given by the linear transpose of \(f\),
+thus \((\partial f(x))^{T}(dy) = f^T(dy)\). Also, all higher order
 derivatives can be expressed in terms of \(f\) and its transpose. To
@@
 Currently, \texttt{JAXbind} only supports primitives that act on CPU
-memory. In the future, GPU support could be added analogous to the CPU
-backend. The automatic differentiation in JAX is backend agnostic and
-would thus not require any additional bindings to work on the GPU.
+memory. In the future, GPU support could be added, which should work
+analogously to the CPU support in most respects. The automatic
+differentiation in JAX is backend agnostic and would thus not require
+any additional bindings to work on the GPU.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-tensorflow2015}{}%
+Abadi, M., Agarwal, A., Barham, P., Brevdo, E., Chen, Z., Citro, C.,
+Corrado, G. S., Davis, A., Dean, J., Devin, M., Ghemawat, S.,
+Goodfellow, I., Harp, A., Irving, G., Isard, M., Jia, Y., Jozefowicz,
+R., Kaiser, L., Kudlur, M., \ldots{} Zheng, X. (2015).
+\emph{{TensorFlow}: Large-scale machine learning on heterogeneous
+systems}. \url{https://www.tensorflow.org/}
+
+\leavevmode\hypertarget{ref-PyTorch2024}{}%
+Ansel, J., Yang, E., He, H., Gimelshein, N., Jain, A., Voznesensky, M.,
+Bao, B., Bell, P., Berard, D., Burovski, E., Chauhan, G., Chourdia, A.,
+Constable, W., Desmaison, A., DeVito, Z., Ellison, E., Feng, W., Gong,
+J., Gschwind, M., \ldots{} Chintala, S. (2024, April). {PyTorch} 2:
+Faster machine learning through dynamic {P}ython bytecode transformation
+and graph compilation. \emph{29th ACM International Conference on
+Architectural Support for Programming Languages and Operating Systems,
+Volume 2 (ASPLOS '24)}. \url{https://doi.org/10.1145/3620665.3640366}
+
 \leavevmode\hypertarget{ref-Resolve2024}{}%
@@
 Scheel-Platz, L. I., Guardiani, M., Eberle, V., Westerkamp, M., \&
-Enßlin, T. A. (2024). \emph{{Re-Envisioning Numerical Information Field
-Theory (NIFTy.re): A Library for Gaussian Processes and Variational
-Inference}}. \url{https://doi.org/10.48550/arXiv.2402.16683}
-
-\leavevmode\hypertarget{ref-Moses2024}{}%
-EnzymeAD. (2024). \emph{{Enzyme-JAX}} (Version 0.0.6) {[}Computer
-software{]}. \url{https://github.com/EnzymeAD/Enzyme-JAX}
+Enßlin, T. A. (2024). Re-envisioning numerical information field theory
+({NIFTy.re}): A library for {G}aussian processes and variational
+inference. \emph{Journal of Open Source Software}, \emph{9}(98), 6593.
+\url{https://doi.org/10.21105/joss.06593}
 
@@
 
+\leavevmode\hypertarget{ref-Moses2024}{}%
+Moses, W. S., \& Zinenko, O. (2024). \emph{{Enzyme-JAX}} (Version 0.0.6)
+{[}Computer software{]}. \url{https://github.com/EnzymeAD/Enzyme-JAX}
+
 \leavevmode\hypertarget{ref-ducc0}{}%

### 2024-06-27 paper/paper.tex

@@
 
-The use of JAX (Bradbury et al., 2018) is widespread in the natural
-sciences. Of particular interest is JAX's powerful transformation
-system. It enables a user to retrieve arbitrary derivatives of
-functions, batch computations, and just-in-time compile code for
-additional performance. Its transformation system requires that all
-components of the computation are written in JAX.
+The use of JAX (\protect\hyperlink{ref-Jax2018}{Bradbury et al., 2018})
+is widespread in the natural sciences. Of particular interest is JAX's
+powerful transformation system. It enables a user to retrieve arbitrary
+derivatives of functions, batch computations, and just-in-time compile
+code for additional performance. Its transformation system requires that
+all components of the computation are written in JAX.
 
@@
 We intend to use this package to connect the Hartley transform and the
-spherical harmonic transform from DUCC (Reinecke, 2024) to the
-probabilistic programming package NIFTy (Edenhofer et al., 2024) as well
-as the radio interferometry response from DUCC with the radio astronomy
-package \texttt{resolve} (Arras et al., 2024). Furthermore, we intend to
-connect the non-uniform FFT from DUCC with JAX for applications in
-strong-lensing astrophysics. We envision many further applications
-within and outside of astrophysics.
+spherical harmonic transform from DUCC
+(\protect\hyperlink{ref-ducc0}{Reinecke, 2024}) to the probabilistic
+programming package NIFTy
+(\protect\hyperlink{ref-Edenhofer2023NIFTyRE}{Edenhofer et al., 2024})
+as well as the radio interferometry response from DUCC with the radio
+astronomy package \texttt{resolve}
+(\protect\hyperlink{ref-Resolve2024}{Arras et al., 2024}). Furthermore,
+we intend to connect the non-uniform FFT from DUCC with JAX for
+applications in strong-lensing astrophysics. We envision many further
+applications within and outside of astrophysics.
 
@@
 the need for C++ or LLVM. The package that comes the closest is
-Enzyme-JAX (W. S. Moses \& Zinenko, 2024), which allows one to bind
-arbitrary LLVM/MLIR, including C++, with automatically-generated (W.
-Moses \& Churavy, 2020; W. S. Moses et al., 2021, 2022) or
+Enzyme-JAX (\protect\hyperlink{ref-Moses2024}{W. S. Moses \& Zinenko,
+2024}), which allows one to bind arbitrary LLVM/MLIR, including C++,
+with automatically-generated (\protect\hyperlink{ref-Moses2021}{W. S.
+Moses et al., 2021}, \protect\hyperlink{ref-Moses2022}{2022};
+\protect\hyperlink{ref-Moses2020}{W. Moses \& Churavy, 2020}) or
 manually-defined derivatives to JAX.
 
-PyTorch (Ansel et al., 2024) and TensorFlow (Abadi et al., 2015) also
-provide interfaces for custom extensions. PyTorch has an extensively
-documented Python interface\footnote{\url{https://pytorch.org/docs/stable/notes/extending.html}}
+PyTorch (\protect\hyperlink{ref-PyTorch2024}{Ansel et al., 2024}) and
+TensorFlow (\protect\hyperlink{ref-tensorflow2015}{Abadi et al., 2015})
+also provide interfaces for custom extensions. PyTorch has an
+extensively documented Python interface\footnote{\url{https://pytorch.org/docs/stable/notes/extending.html}}
 for wrapping custom Python functions as PyTorch functions. This
@@
 Goodfellow, I., Harp, A., Irving, G., Isard, M., Jia, Y., Jozefowicz,
-R., Kaiser, L., Kudlur, M., \ldots{} Zheng, X. (2015).
-\emph{{TensorFlow}: Large-scale machine learning on heterogeneous
-systems}. \url{https://www.tensorflow.org/}
+R., Kaiser, L., Kudlur, M., … Zheng, X. (2015). \emph{{TensorFlow}:
+Large-scale machine learning on heterogeneous systems}.
+\url{https://www.tensorflow.org/}
 
@@
 Constable, W., Desmaison, A., DeVito, Z., Ellison, E., Feng, W., Gong,
-J., Gschwind, M., \ldots{} Chintala, S. (2024, April). {PyTorch} 2:
-Faster machine learning through dynamic {P}ython bytecode transformation
-and graph compilation. \emph{29th ACM International Conference on
+J., Gschwind, M., … Chintala, S. (2024, April). {PyTorch} 2: Faster
+machine learning through dynamic {P}ython bytecode transformation and
+graph compilation. \emph{29th ACM International Conference on
 Architectural Support for Programming Languages and Operating Systems,
-Volume 2 (ASPLOS '24)}. \url{https://doi.org/10.1145/3620665.3640366}
+Volume 2 (ASPLOS ’24)}. \url{https://doi.org/10.1145/3620665.3640366}
 
@@
 S., \& Zhang, Q. (2018). \emph{{JAX}: Composable transformations of
-{P}ython+{N}um{P}y programs} (Version 0.3.13) {[}Computer software{]}.
+{P}ython+{N}um{P}y programs} (Version 0.3.13).
 \url{http://github.com/google/jax}
@@
 \url{https://doi.org/10.21105/joss.06593}
-
-\leavevmode\hypertarget{ref-Moses2020}{}%
-Moses, W., \& Churavy, V. (2020). Instead of rewriting foreign code for
-machine learning, automatically synthesize fast gradients. In H.
-Larochelle, M. Ranzato, R. Hadsell, M. F. Balcan, \& H. Lin (Eds.),
-\emph{Advances in neural information processing systems} (Vol. 33, pp.
-12472--12485). Curran Associates, Inc.
-\url{https://proceedings.neurips.cc/paper/2020/file/9332c513ef44b682e9347822c2e457ac-Paper.pdf}
 
@@
 \leavevmode\hypertarget{ref-Moses2024}{}%
-Moses, W. S., \& Zinenko, O. (2024). \emph{{Enzyme-JAX}} (Version 0.0.6)
-{[}Computer software{]}. \url{https://github.com/EnzymeAD/Enzyme-JAX}
+Moses, W. S., \& Zinenko, O. (2024). \emph{{Enzyme-JAX}} (Version
+0.0.6). \url{https://github.com/EnzymeAD/Enzyme-JAX}
+
+\leavevmode\hypertarget{ref-Moses2020}{}%
+Moses, W., \& Churavy, V. (2020). Instead of rewriting foreign code for
+machine learning, automatically synthesize fast gradients. In H.
+Larochelle, M. Ranzato, R. Hadsell, M. F. Balcan, \& H. Lin (Eds.),
+\emph{Advances in neural information processing systems} (Vol. 33, pp.
+12472–12485). Curran Associates, Inc.
+\url{https://proceedings.neurips.cc/paper/2020/file/9332c513ef44b682e9347822c2e457ac-Paper.pdf}
 
@@
 Reinecke, M. (2024). \emph{{DUCC}: Distinctly useful code collection}
-(Version 0.33.0) {[}Computer software{]}.
-\url{https://gitlab.mpcdf.mpg.de/mtr/ducc}
+(Version 0.33.0). \url{https://gitlab.mpcdf.mpg.de/mtr/ducc}
 
